---
title: "Null Hypothesis"
author: "Sophie Wulfing"
date: "2/12/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(tidyverse)

setwd("~/STOICH.Aim1/Code/SWulfing")


ggplotRegression <- function (fit) {
  
  require(ggplot2)
  
  ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
    geom_point() +
    stat_smooth(method = "lm", col = "red") +
    labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                       "Intercept =",signif(fit$coef[[1]],5 ),
                       " Slope =",signif(fit$coef[[2]], 5),
                       " P =",signif(summary(fit)$coef[2,4], 5)))
}

test_data <- data.frame(read.csv("ChemData_NLArs2013.csv"))
```

## Note

This is what I've come up with regarding how to tackle the null hypothesis. I know we've taken NLA data out of circulation but I'm going to continue to use it here only because I was using it earlier as an exampel set and I wanted to get feedback from you all about the methods and learn a little more about the changes to data before I applied it to other datasets.

```{r data}
model_2013 <- lm(NITRATE_N_RESULT~DOC_RESULT, data = test_data)
summary(model_2013)

test_dataLOGS <- test_data %>%
  mutate(log_Nitrate = log(NITRATE_N_RESULT)) %>%
  mutate(log_TP = log(PTL_RESULT)) %>%
  mutate(log_DOC = log(DOC_RESULT))
                                   # Duplicate data
test_dataLOGS$log_Nitrate[is.na(test_dataLOGS$log_Nitrate) | test_dataLOGS$log_Nitrate == "-Inf"] <- NA

quantile(test_dataLOGS$DOC_RESULT)

q4 <- quantile(test_data$DOC_RESULT)[4]
q5 <- quantile(test_data$DOC_RESULT)[5]

lowDOC_data <- test_data %>%
  filter(DOC_RESULT < quantile(test_data$DOC_RESULT)[4])

par(mfrow = c(1,2))
p <- ggplot(test_dataLOGS) +
  geom_point(aes(DOC_RESULT, NITRATE_N_RESULT, color = PTL_RESULT * 1000)) +
  scale_color_viridis_c("TP"~(mu~g~L^-1)) +
  theme_bw() +
  labs(title = "NRSA 2013",
       x = "DOC", y = "NITRATE")
p + geom_vline(aes(xintercept = q4)) + geom_vline(aes(xintercept = q5))


ggplot(lowDOC_data) +
  geom_point(aes(DOC_RESULT, NITRATE_N_RESULT, color = PTL_RESULT * 1000)) +
  scale_color_viridis_c("TP"~(mu~g~L^-1)) +
  theme_bw() +
  labs(title = "NRSA 2013",
       x = "DOC", y = "NITRATE")
```

## Note:
For now I am using the 4the quantile of this specific dataset. I realize that will change both where the quantile is/where we choose to cut off data


```{r to log or not to log, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
