---
title: "EU Dataset tidying"
author: "Daniel Gschwenter"
date: "3/9/2022 Updated at `r Sys.time() `"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Packages.
lapply(c("tidyverse", "glue", "knitr", "kableExtra"), require, character.only = T)

```

---
---

### Code to tidy EU data set following GitHub Guidelines

---
---

#### Guidelines
---
###### Major ions to include

- CHLORIDE
- SULFATE
- CALCIUM
- MAGNESIUM
- SODIUM
- POTASSIUM
- IRON
- MANGANESE

##### Forms of C,N,P

Naming conventions

- Total kjeldahl nitrogen == TKN
- Total particulate phosphorus == TPP
- Total carbon == TC
- NH4 == NH4 as N
- NH3 == NH3 as N
- NO2 == NO2 as N
- NO3 == NO3 as N

Conversion of N forms to N as N: Divide mass of species by atomic weight and multiply by atomic weight of N.

```{r N conversion table, include=F}

N <- tibble(
  Species = c("NH4", "NH3", "NO3", "NO2"), 
  "Atomic weight" = c(18.04, 17.031, 62.0049, 46.0055), 
  "Atomic weight N" = 14.0067 
)

```

The formula for the conversions of N species are as follows: `r glue("{NSpecies} / {AtomWeight} * {AtomWeightN}", NSpecies = N$Species, AtomWeight = N$"Atomic weight", AtomWeightN = N$"Atomic weight N")`.

DOC, NO3 as N and PO4 as P are already standardised.


##### Other things

- TEMP = deg_C
- SP_COND = uS/cm
- DO = mg/L
- CHLA = mg/L
- pH

##### Data upload.

When you are finished, save the new dataset as "NAME_1.csv" in the Data/other_vars_datasets folder, i.e., do not overwrite the dataset that is in there, rather make a new one, e.g., EU_1.csv.

---


```{r Loading data, include=F, warning=F, message=F}

#dat <- read_csv("https://raw.githubusercontent.com/swulfing/STOICH.Aim1/main/Data/other_vars_datasets/EU.csv")

# Save data locally for easier reading.

#if(!dir.exists("local_data")){dir.create("local_data")}
#write_delim(dat, "local_data/EU_all_raw.csv", delim = ",")
dat <- read_csv("C://Users/dgschwentner2/OneDrive - University of Nebraska-Lincoln/PhD/Aim1Biogeochem/local_data/EU_all_raw.csv")

```


```{r Check out columns etc., echo = F, warning=F, message=F}

head(dat) %>%
  kable() %>%
  kable_classic(full_width = T)


summary(dat) %>% 
  kable() %>%
  kable_classic(full_width = T)
  

```

```{r Check out data  types and their units, echo = F}

# Copy of data set.
dat_edits <- dat %>%
  mutate(param_units = paste0(VARIABLE, " ", UNIT))

### Do not run, this just creates a headache.
# dat_edits %>%
#   group_by(param_units) %>%
#   mutate(n = n()) %>%
#   kable() %>%
#   kable_classic(full_width = F, position = "float_left")

# Do not run.
#sort(unique(dat_edits$param_units))


```

There are a total of `r length(unique(dat_edits$param_units))` unique variables and units in the data set. YIKES!

```{r Create overview of desired data, include=F}

# Filter for required units.

want <-tibble(
  Param= c(
  "Calcium mg/L" , "Chloride mg/L", "Chlorophyll a ug/L" , "DOC mg/L" , "Electrical conductivity uS/cm" , 
  "Iron and its compounds ug/L", "Kjeldahl nitrogen mg{N}/L", "Magnesium mg/L", "Manganese and its compounds ug/L", 
  "Nitrite mg{NO2}/L", "NO3 as N mg/L", "Non-ionised ammonia mg{NH3}/L", "Ammonium mg{NH4}/L" , "PO4 as P mg/L", "pH [pH]",
  "Potassium mg/L", "Sodium mg/L", "Sulphate mg/L", "Total inorganic nitrogen mg{N}/L" , "Total oxidised nitrogen mg{N}/L", 
  "Total nitrogen mg{N}/L", "Total organic nitrogen mg{N}/L", "Total organic carbon (TOC) mg{C}/L", "Total phosphorus mg{P}/L",
  "Water temperature Cel" , "Dissolved oxygen mg/L"
), 
Param_new = c(
  "CALCIUM", "CHLORIDE", "CHLA", "DOC", "SP_COND", "IRON", "TKN", "MAGNESIUM", 
  "MANGANESE", "NO2 as N", "NO3 as N", "NH3 as N", "NH4 as N", "PO4 as P", "pH", "POTASSIUM", 
  "SODIUM", "SULPHATE", "TIN as N", "TOXIDN as N", "TN", "TON as N", "TOC", "TP", 
  "TEMP", "DO"
),
Unit = c("mg/L", "mg/L", "mg/L", "mg/L", "uS/cm", "mg/L",  "mg/L", "mg/L", "mg/L", "mg/L", "mg/L", "mg/L", "mg/L", "mg/L", "Unitless", 
         "mg/L", "mg/L", "mg/L", "mg/L", "mg/L", "mg/L", "mg/L", "mg/L", "mg/L", "deg_C", "mg/L"), 
Action = c("Rename", "Rename", "Rename and /1000", "Rename", "Rename", "Rename and /1000", "Rename", "Rename", 
           "Rename and /1000", "Rename and /46.0*14.0", "Rename and /62.0*14.0", "Rename and /17.0*14.0", "Rename and /18.0*14.0",
           "Rename", "Rename", "Rename", "Rename", "Rename", "Rename", "Rename", "Rename", "Rename", "Rename", "Rename", "Rename", "Rename"))


```

In total, there are `r length(want)` parameters that should be extracted from the data set. A list of the parameters and their renamed and converted names is presented below for quality control. `r want %>% kable(caption = "Required parameters and naming conventions") %>% kable_classic(full_width = F)`


```{r Filter data set, include=F}

# Filter.

# Create a test data set to compare that conversions were in fact gucci.

test_filter <- data_filtered <- dat_edits %>%
  filter(param_units %in% want$Param) %>%
  merge(want, by.x = "param_units", by.y = "Param") %>% # Some clever merging to make renaming effortless....
  select( # Potentially not the best way to do it.
    DATE_COL, SITE_ID, "VARIABLE" = Param_new, RESULT, "UNIT" = Unit, LAT, LON, ECO_TYPE 
  )

# Filter actual data set.
data_filtered <- dat_edits %>%
  filter(param_units %in% want$Param) %>%
  merge(want, by.x = "param_units", by.y = "Param") %>% # Some clever merging to make renaming effortless....
  select( # Potentially not the best way to do it.
    DATE_COL, SITE_ID, "VARIABLE" = Param_new, RESULT, "UNIT" = Unit, LAT, LON, ECO_TYPE 
  ) %>%
  mutate( # I don't think this is the best or safest way to perform the conversions; happy to do it manually if people want a safer approach.
    RESULT = ifelse(VARIABLE == "CHLA", RESULT/1000, RESULT), 
    RESULT = ifelse(VARIABLE == "IRON", RESULT/1000, RESULT), 
    RESULT = ifelse(VARIABLE == "MANGANESE", RESULT/1000, RESULT), 
    RESULT = ifelse(VARIABLE == "NO2 as N", RESULT/46.0055*14.0067, RESULT),
    RESULT = ifelse(VARIABLE == "NO3 as N", RESULT/62.004*14.0067, RESULT),
    RESULT = ifelse(VARIABLE == "NH3 as N", RESULT/17.031*14.0067, RESULT),
    RESULT = ifelse(VARIABLE == "NH4 as N", RESULT/18.04*14.0067,RESULT)
    
)


```


```{r Some quality control plots, include=F}

edited_vars <- c("CHLA", "IRON", "MANGANESE", "NO2 as N", "NO3 as N", "NH3 as N", "NH4 as N")

for(i in edited_vars){
  plot(test_filter[test_filter$VARIABLE == i,"RESULT"], data_filtered[data_filtered$VARIABLE ==i,"RESULT"], main = i)
}

# Seemed to work well...

```

```{r Save data, include=F}

eu_edited <- data_filtered

#write_delim(eu_edited, "C://Users/dgschwentner2/OneDrive - University of Nebraska-Lincoln/PhD/Aim1Biogeochem/local_data/EU_filtered_cleaned.csv", delim = "'")


```


An overview of the edited data set is presented in the tables below.

```{r Overview of filtered data, echo = F}

eu_edited %>%
  mutate(VARIABLE = paste0(VARIABLE, " ", UNIT)) %>%
  group_by(VARIABLE) %>%
  summarise(
    Min = min(RESULT),
    Mean = mean(RESULT), 
    Max = max(RESULT), 
    "CV in %" = sd(RESULT)/mean(RESULT) * 100,
    "25th percentile" = quantile(RESULT, 0.25),
    "75th percentile" = quantile(RESULT, 0.75)
  ) %>%
  kable(digits = 2) %>%
  kable_classic()


eu_edited %>%
  mutate(VARIABLE = paste0(VARIABLE, " ", UNIT)) %>%
  group_by(VARIABLE, ECO_TYPE) %>%
  summarise(
    Min = min(RESULT),
    Mean = mean(RESULT), 
    Max = max(RESULT), 
    "CV in %" = sd(RESULT)/mean(RESULT) * 100,
    "25th percentile" = quantile(RESULT, 0.25),
    "75th percentile" = quantile(RESULT, 0.75)
  ) %>%
  kable(digits = 2) %>%
  kable_classic()


```

