---
title: "NRC Data Set Cleaning"
author: "Daniel Gschwenter"
date: "3/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

---
---

### Code to tidy NRC data set following GitHub Guidelines

---
---

#### Guidelines
---
###### Major ions to include

- CHLORIDE
- SULFATE
- CALCIUM
- MAGNESIUM
- SODIUM
- POTASSIUM
- IRON
- MANGANESE

##### Forms of C,N,P

Naming conventions

- Total kjeldahl nitrogen == TKN
- Total particulate phosphorus == TPP
- Total carbon == TC
- NH4 == NH4 as N
- NH3 == NH3 as N
- NO2 == NO2 as N
- NO3 == NO3 as N

Conversion of N forms to N as N: Divide mass of species by atomic weight and multiply by atomic weight of N.

```{r N conversion table, include=F}

N <- tibble(
  Species = c("NH4", "NH3", "NO3", "NO2"), 
  "Atomic weight" = c(18.04, 17.031, 62.0049, 46.0055), 
  "Atomic weight N" = 14.0067 
)

```

The formula for the conversions of N species are as follows: `r glue("{NSpecies} / {AtomWeight} * {AtomWeightN}", NSpecies = N$Species, AtomWeight = N$"Atomic weight", AtomWeightN = N$"Atomic weight N")`. Note: this is not needed for NO~3~^-^ as already converted to NO~3~^-^ as N.

DOC, NO3 as N and PO4 as P are already standardised.


##### Other things

- TEMP = deg_C
- SP_COND = uS/cm
- DO = mg/L
- CHLA = mg/L
- pH

##### Data upload.

When you are finished, save the new dataset as "NAME_1.csv" in the Data/other_vars_datasets folder, i.e., do not overwrite the dataset that is in there, rather make a new one, e.g., EU_1.csv.

---

```{r Load data save locally and re-import, include=F}

# nrc <- read_csv("https://raw.githubusercontent.com/swulfing/STOICH.Aim1/main/Data/other_vars_datasets/NRC.csv")
# write_delim(nrc, "local_data/nrc_raw.csv", delim = ",")
nrc <- read_csv("local_data/nrc_raw.csv")

# Had to manually replace µ with u in data set...
```


```{r Overview of data, echo = F}

colnames(nrc)

unique(nrc$VARIABLE) # Yikes!!!!

# Copy of data set.
nrc_edit<- nrc %>%
  mutate(param_units = paste0(VARIABLE, " ", UNIT))

### Do not run, this just creates a headache.
# dat_edits %>%
#   group_by(param_units) %>%
#   mutate(n = n()) %>%
#   kable() %>%
#   kable_classic(full_width = F, position = "float_left")

# Do not run.
sort(unique(nrc_edit$param_units))


```

N.B.!!!! For the NRC data set following parameters have issues:
- Calcium: doesn't exist in data set.
- Chloride: only dissolved fraction present.
- Sodium: only dissolved fraction present.
- Potassium: only dissolved fraction present.
- Magneisum is the sum of the particulate and dissolved fractions.
- Assuming following vars are already converted to N based on names and prior information: AMMONIA DISSOLVED MG/L N, 
- Particulate P calculated as TP - Dissolved P

Lots of dissolved N that I did not include in data tidying.


```{r ID variables of interest, echo = F}

# Variables of interest:
vars <- c("CARBON TOTAL ORGANIC MG/L", # Transform to TC as C.
          "CHLORIDE DISSOLVED MG/L", # Rename
          "IRON TOTAL uG/L" , # ug/L to mg/L -> divide by 1000.
          "MAGNESIUM DISSOLVED MG/L",  # Rename
          "MAGNESIUM PARTICULATE uG/G",# Rename, convert to mg/L and sum with dissolved magnesium
          "MANGANESE TOTAL uG/L", # Convert to mg/L
          "POTASSIUM DISSOLVED MG/L", # Rename
          "SODIUM DISSOLVED/FILTERED MG/L", # Rename
          "AMMONIA DISSOLVED MG/L N", # Rename to NH3
          "CHLOROPHYLL A uG/L", # Convert to mg/L
          "CONDUCTIVITY uS/CM" , # Rename
          "DOC MG/L", # Rename
          "NITROGEN NITRITE MG/L", # Convert to NO2 as N
          "NO3 as N MG/L", # Rename
          "NITROGEN TOTAL MG/L", # Convert to TN as N
          "PH (FIELD) PH", # Rename
          "PHOSPHORUS TOTAL DISSOLVED MG/L P", # Rename
          "PHOSPHORUS TOTAL MG/L P", # Rename and use to calculate particulate P from dissolved P
          "PO4 as P MG/L P", # Rename
          "TEMPERATURE WATER (FIELD) DEG C", # Rename
          "OXYGEN DISSOLVED (FIELD) MG/L"  # Rename
)


# Filter.
nrc_select <- nrc_edit %>%
  filter(param_units %in%  vars) %>%
  pivot_wider(id_cols = c(DATE_COL, SITE_ID, LAT, LON, ECO_TYPE), # Pivoting makes things a bit more managable for what we need to do.
              names_from = param_units, 
              values_from = RESULT,
              values_fn = mean) # Some list cols so had to summarise.

```


```{r Data editing, include=F}


nrc_clean <- nrc_select %>%
  mutate("TCasC mg/L" = `CARBON TOTAL ORGANIC MG/L`/12.011, # Divide by molecular weight of C
         "CHLORIDE mg/L" = `CHLORIDE DISSOLVED MG/L`,
         "IRON mg/L" = `IRON TOTAL uG/L`/1000,
         "MAGNESIUM mg/L" = `MAGNESIUM DISSOLVED MG/L` + `MAGNESIUM PARTICULATE uG/G`/1000,
         "MANGANESE mg/L" = `MANGANESE TOTAL uG/L`/1000,
         "POTASSIUM mg/L" = `POTASSIUM DISSOLVED MG/L`, 
         "SODIUM mg/L" = `SODIUM DISSOLVED/FILTERED MG/L`,
         "NH3asN mg/L" = `AMMONIA DISSOLVED MG/L N`,
         "CHLA mg/L" = `CHLOROPHYLL A uG/L`/1000,
         "SP_Cond uS/cm" = `CONDUCTIVITY uS/CM`,
         "DOC mg/L" = `DOC MG/L`,
         "NO2asN mg/L" = `NITROGEN NITRITE MG/L`/46.0055,
         "TNasN mg/L" = `NITROGEN TOTAL MG/L`/14.007, 
         "pH Unitless" = `PH (FIELD) PH`,
         "TDPasP mg/L" = `PHOSPHORUS TOTAL DISSOLVED MG/L P`,
         "TPasP mg/L" = `PHOSPHORUS TOTAL MG/L P`,
         "TPPasP mg/L" = `TPasP mg/L` - `TDPasP mg/L`,
         "PO4asP mg/L" = `PO4 as P MG/L P`,
         "TEMP deg_C" = `TEMPERATURE WATER (FIELD) DEG C`,
         "DO mg/L" = `OXYGEN DISSOLVED (FIELD) MG/L`
         ) %>%
  select(
    DATE_COL, SITE_ID, LAT, LON, ECO_TYPE, "TCasC mg/L", "CHLORIDE mg/L", "IRON mg/L", "MAGNESIUM mg/L", "MANGANESE mg/L", "POTASSIUM mg/L",
    "SODIUM mg/L", "NH3asN mg/L", "CHLA mg/L", "SP_Cond uS/cm", "DOC mg/L", "NO2asN mg/L", "TNasN mg/L", "pH Unitless", "TDPasP mg/L",
    "TPasP mg/L", "TPPasP mg/L", "PO4asP mg/L", "TEMP deg_C", "DO mg/L"
  )

# Pivot back to long format...

nrc_clean_long <- nrc_clean %>%
  gather("VARIABLE", "RESULT", -DATE_COL, -SITE_ID, -LAT, -LON, -ECO_TYPE) %>%
  separate(VARIABLE, c("VARIABLE", "UNIT"), sep = " ") %>%
  drop_na() %>%
  mutate( # Final renaming.
    VARIABLE = ifelse(VARIABLE == "TCasC","TC as C", VARIABLE),
    VARIABLE = ifelse(VARIABLE == "NH3asN", "NH3 as N", VARIABLE),
    VARIABLE = ifelse(VARIABLE == "NO2asN", "NO2 as N", VARIABLE),
    VARIABLE = ifelse(VARIABLE == "TNasN", "TN as N", VARIABLE),
    VARIABLE = ifelse(VARIABLE == "TDPasP", "TDP as P", VARIABLE),
    VARIABLE = ifelse(VARIABLE == "TPasP", "TP as P", VARIABLE),
    VARIABLE = ifelse(VARIABLE == "TPPasP", "TPP as P", VARIABLE),
    VARIABLE = ifelse(VARIABLE == "PO4asP", "PO4 as P", VARIABLE),
    UNIT = ifelse(UNIT == "Unitless", NA, UNIT)
  )

# Quality control

sort(unique(paste0(nrc_clean_long$VARIABLE, " ", nrc_clean_long$UNIT)))

```

```{r Save file}

write_delim(nrc_clean_long, "local_data/nrc_cleaned.csv", delim = ",")

```

