---
title: "Aim 1 EDA Vis"
author: "Daniel Gschwentner"
date: "24 March 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

lapply(c("tidyverse", "ggsci", "ggpubr"), require, character.only = T)

my_theme <- theme_bw() + 
  theme(axis.text = element_text(color = "black"), 
        axis.ticks = element_line(color = "black"), 
        panel.grid = element_blank(),
        strip.background = element_rect(size= 0.75, fill = "white"),
        legend.position = "bottom")

theme_set(my_theme)


```


```{r Load data, include=F}

source("C://Users/dgschwentner2/OneDrive - University of Nebraska-Lincoln/PhD/Aim1Biogeochem/R Code/LRock_AllMasterVars.R")

```

```{r Pivoting, include=F}

# Pivot to include all vars for plotting.

master_wide <- ALL_CNP_VARS %>%
  pivot_wider(id_cols = c(DATE_COL, SITE_ID,  LAT, LON, ECO_TYPE), # Gotta drop units here.
              names_from = VARIABLE, values_from = RESULT, values_fn = mean)
# This is a bit of an issue....
# Long data format hides duplicate measurments from same site, date.... Not good...

# Some maniputlation so I can get subsets for vis purposes
CNP_subsets <- ALL_CNP_VARS %>%
  mutate(SUBSET = "ALL") %>%
  bind_rows(ALL_CNP_VARS %>% filter(ECO_TYPE == "River/Stream") %>% mutate(SUBSET = "River/Stream")) %>%
  bind_rows(ALL_CNP_VARS %>% filter(ECO_TYPE != "River/Stream") %>% mutate(SUBSET = "Lake"))


```

```{r Boxplots, echo = F, fig.align='center', dpi = 600}

# Density distribution of variables.
# Gott use log, not really feasible to see stuff otherwise....

(box_all <- CNP_subsets %>%
  ggplot(aes(SUBSET, log10(RESULT), fill = SUBSET)) + 
  geom_boxplot() +
  scale_fill_jama() + 
  scale_color_jama() + 
  facet_wrap(VARIABLE~.,scales = "free") + 
  labs(x = "Variable in mg/L, deg C, pH units or uS/Cm", y = "Density"))


boxplt_list <- list()
for(i in unique(CNP_subsets$VARIABLE)){
  boxplt_list[[i]] <- CNP_subsets %>%
    filter(VARIABLE == i) %>%
  ggplot(aes(SUBSET, RESULT, fill = SUBSET)) + 
  geom_boxplot() +
  scale_fill_jama() + 
  scale_color_jama() + 
  labs(y = "unit", x = NULL, subtitle = i)

}

boxplt_list_log10 <- list()
for(i in unique(CNP_subsets$VARIABLE)){
  boxplt_list_log10[[i]] <- CNP_subsets %>%
    filter(VARIABLE == i) %>%
  ggplot(aes(SUBSET, log10(RESULT), fill = SUBSET)) + 
  geom_boxplot() +
  scale_fill_jama() + 
  scale_color_jama() + 
  labs(y = "log10(unit)", x = NULL, subtitle = i)

}

# A subset of the data; just the most relevant aspects....

ggarrange(plotlist = boxplt_list[c("DOC", "NO3 as N", "PO4 as P", "TN", "TP" ,"pH", "DO", "SP_COND")], ncol = 3, nrow = 3, common.legend = T, legend = "bottom")


ggarrange(plotlist =boxplt_list_log10[c("DOC", "NO3 as N", "PO4 as P", "TN", "TP" ,"pH", "DO", "SP_COND")], ncol = 3, nrow = 3, common.legend = T, legend = "bottom")

```

---

Some thoughts:

- Rivers/Streams generally have higher NO~3~ as N concentrations than lakes. Average NO~3~ as N concentrations for rivers and streams are `r mean(CNP_subsets %>% filter(SUBSET == "River/Stream" & VARIABLE == "NO3 as N") %>% pull(RESULT)) ` and `r mean(CNP_subsets %>% filter(SUBSET == "Lake" & VARIABLE == "NO3 as N") %>% pull(RESULT)) ` mgL. 

- PO~4~ concentrations are similarily lower in lakes than in rivers or streams. Average concentrations are `r mean(CNP_subsets %>% filter(SUBSET == "Lake" & VARIABLE == "PO4 as P") %>% pull(RESULT)) ` and `r  mean(CNP_subsets %>% filter(SUBSET == "River/Stream" & VARIABLE == "PO4 as P") %>% pull(RESULT))` mgL respectively.

- TP and TN pools are similar across lakes and rivers/streams.

- pH and specific conductivity are similar, however lakes have a larger range of pH. Both ecotypes have more neutral and basic pH than acidic. 

---

```{r Stoich ratios boxplot, echo=F}

# First calc. molar concentrations.

stoic_ratios <- ALL_CNP %>%
  mutate("DOC as C molar" = DOC/1000/12.011,
         "NO3 as N molar" = NO3.as.N/1000/14.007, 
         "PO4 as P molar"= PO4.as.P/1000/30.974,
         "DOC:NO3" = `DOC as C molar`/`NO3 as N molar` ,
         "DOC:PO4" = `DOC as C molar`/`PO4 as P molar`,
         "NO3:PO4" = `NO3 as N molar`/`PO4 as P molar`,
         SUBSET = "ALL")
  
stoic_ratios <- stoic_ratios %>%
  bind_rows(stoic_ratios %>% filter(ECO_TYPE == "River/Stream") %>% mutate(SUBSET ="River/Stream")) %>%
  bind_rows(stoic_ratios %>% filter(ECO_TYPE == "Lake") %>% mutate(SUBSET = "Lake"))

ggarrange(
  
  stoic_ratios %>% ggplot(aes(SUBSET,`DOC:NO3`, fill = SUBSET )) + 
    geom_boxplot() + geom_hline(yintercept = 106/16, col = "red") + 
    scale_fill_jama() + labs(y = "DOC:NO3", x = NULL),
  
  stoic_ratios %>% ggplot(aes(SUBSET,`DOC:PO4`, fill = SUBSET )) + 
    geom_boxplot() + geom_hline(yintercept = 106/1, col = "red") + 
    scale_fill_jama() + labs(y = "DOC:PO4", x = NULL),
  
  stoic_ratios %>% ggplot(aes(SUBSET,`NO3:PO4`, fill = SUBSET )) + 
    geom_boxplot() + geom_hline(yintercept = 16/1, col = "red") + 
    scale_fill_jama() + labs(y = "NO3:PO4", x = NULL),
  
  stoic_ratios %>% ggplot(aes(SUBSET,log10(`DOC:NO3`), fill = SUBSET )) + 
    geom_boxplot() + geom_hline(yintercept = log10(106/16), col = "red") + 
    scale_fill_jama() + labs(y = "log10(DOC:NO3)", x = NULL),
  
  stoic_ratios %>% ggplot(aes(SUBSET,log10(`DOC:PO4`), fill = SUBSET )) + 
    geom_boxplot() + geom_hline(yintercept = log10(106/1), col = "red") + 
    scale_fill_jama() + labs(y = "DOC:PO4", x = NULL),
  
  stoic_ratios %>% ggplot(aes(SUBSET,log10(`NO3:PO4`), fill = SUBSET )) + 
    geom_boxplot() + geom_hline(yintercept = log10(16/1), col = "red") + 
    scale_fill_jama() + labs(y = "log10(NO3:PO4)", x = NULL), 
  
  ncol = 3, nrow = 2, common.legend = T, legend = "bottom"
  
)

```


---

Some comments

- Stoich. imbalances abound: as far from Redfield as you can imagine....

- Lakes and streams have similar NO~3~:PO~4~ but DOC:nutrient ratios are very distinct. Suggests DOC dynamics in lakes and streams are different...

- Average values for all the data are as follows:

`r flextable::flextable(stoic_ratios %>% group_by(SUBSET) %>% summarise_at(c("DOC:NO3", "DOC:PO4", "NO3:PO4"), .funs = function(x){round(median(x), digits = 0)}))`

- Some pretty extreme outliers

---

```{r Density distributions, echo = F, fig.align='center', dpi = 600}

# Density distribution of variables.
density_all <- ALL_CNP_VARS %>%
  mutate(SUBSET = "ALL") %>%
  bind_rows(ALL_CNP_VARS %>% filter(ECO_TYPE == "River/Stream") %>% mutate(SUBSET = "River/Stream")) %>%
  bind_rows(ALL_CNP_VARS %>% filter(ECO_TYPE != "River/Stream") %>% mutate(SUBSET = "Lake")) %>%
  ggplot(aes(RESULT, col = SUBSET)) + 
  geom_density() + 
  scale_fill_jama() + 
  scale_color_jama() + 
  facet_wrap(VARIABLE~.,scales = "free") + 
  labs(x = "Variable in mg/L, deg C, pH units or uS/Cm", y = "Density")

# Log for better vis.
density_alllog10 <- ALL_CNP_VARS %>%
  mutate(SUBSET = "ALL") %>%
  bind_rows(ALL_CNP_VARS %>% filter(ECO_TYPE == "River/Stream") %>% mutate(SUBSET = "River/Stream")) %>%
  bind_rows(ALL_CNP_VARS %>% filter(ECO_TYPE != "River/Stream") %>% mutate(SUBSET = "Lake")) %>%
  ggplot(aes(log10(RESULT), col = SUBSET)) + 
  geom_density() + 
  scale_fill_jama() + 
  scale_color_jama() + 
  facet_wrap(VARIABLE~.,scales = "free") + 
  labs(x = "Variable in log10(mg/L, deg C, pH units or uS/Cm)", y = "Density")

# This is easier vis tbh...

density_list <- list()
for(i in unique(CNP_subsets$VARIABLE)){
 density_list[[i]] <- CNP_subsets %>%
    filter(VARIABLE == i) %>%
  ggplot(aes(RESULT, col = SUBSET, fill = SUBSET, alpha = 0.1)) + 
  geom_density() +
  scale_fill_jama() + 
  scale_color_jama() + 
  labs(y = "Density", x = i, subtitle = i)

}


density_list_log10 <- list()
for(i in unique(CNP_subsets$VARIABLE)){
 density_list_log10[[i]] <- CNP_subsets %>%
    filter(VARIABLE == i) %>%
  ggplot(aes(log10(RESULT), col = SUBSET, fill = SUBSET, alpha = 0.05)) + 
  geom_density() +
  scale_fill_jama() + 
  scale_color_jama() + 
  labs(y = "Density", x = paste0("log10(",i, ")"), subtitle = i)

}
# A subset of the data; just the most relevant aspects....

ggarrange(plotlist = density_list[c("DOC", "NO3 as N", "PO4 as P", "TN", "TP" ,"pH", "DO", "SP_COND")], ncol = 3, nrow = 3, common.legend = T, legend = "bottom")


ggarrange(plotlist =density_list_log10[c("DOC", "NO3 as N", "PO4 as P", "TN", "TP" ,"pH", "DO", "SP_COND")], ncol = 3, nrow = 3, common.legend = T, legend = "bottom")



```

---

Some comments:

For untransformed data:

- Data departs from assumptions of normality (i.e. we need to transform or consider some statistical methods to account for this, e.g. non-parametric approaches)

- Lakes generally have broader range of values: this is the case for DO, temp., pH. Streams and rivers tend to have narrow-ish ranges for DO and temp. and the data suggests that DO and temp. dynamics are different in streams/rivers when compared to lakes. This may need to be considered for analysis purposes.

- Data is from neutral to basic/alkaline systems; not that much data from acidic aquatic ecosystems. 

---

```{r Scatter plots DOC, echo = F}

# Only really interested in relationship between DOC, NO3 and PO4 with other vars...
# Some clever plotting.
scatter_data <- master_wide %>%
  gather("key", "value", -DATE_COL, -SITE_ID, -LAT, -LON, -ECO_TYPE, -DOC, -`NO3 as N`, -`PO4 as P`) 
scatter_data  <- scatter_data %>%
  mutate(SUBSET = "ALL") %>% 
  bind_rows(scatter_data %>% filter(ECO_TYPE == "River/Stream") %>% mutate(SUBSET = "River/Stream")) %>%
  bind_rows(scatter_data %>% filter(ECO_TYPE == "Lake") %>% mutate(SUBSET = "Lake"))

# Scatter lists of plots.
doc_scatter <- list()
for(i in unique(scatter_data$key)){
  doc_scatter[[paste0("DOC X ", i)]] <- scatter_data %>%
  filter(key ==i) %>%
  ggplot(aes(log10(value), log10(DOC), col = SUBSET), alpha = 0.05) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  scale_color_jama() + 
  labs(y = "log10(DOC)", x = paste0("log10(",i,")"), subtitle = paste0("DOC X ", i))
  
}

ggarrange(plotlist = doc_scatter[c("DOC X TN", "DOC X TP" ,"DOC X pH", "DOC X DO", "DOC X SP_COND")], ncol = 3, nrow = 2, common.legend = T, legend = "bottom")



```

```{r Scatter plots NO3 as N, echo = F}

# Only really interested in relationship between DOC, NO3 and PO4 with other vars...
# Some clever plotting.

# Scatter lists of plots.
no3_scatter <- list()
for(i in unique(scatter_data$key)){
  no3_scatter[[paste0("NO3 X ", i)]] <- scatter_data %>%
  filter(key ==i) %>%
  ggplot(aes(log10(value), log10(`NO3 as N`), col = SUBSET)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  scale_color_jama() + 
  labs(y = "log10(NO3)", x = paste0("log10(",i,")"), subtitle = paste0("NO3 X ", i))
  
}

ggarrange(plotlist = no3_scatter[c("NO3 X TN", "NO3 X TP" ,"NO3 X pH", "NO3 X DO", "NO3 X SP_COND")], ncol = 3, nrow = 2, common.legend = T, legend = "bottom")



```

```{r Scatter plots PO4 as P, echo = F}

# Only really interested in relationship between DOC, NO3 and PO4 with other vars...
# Some clever plotting.

# Scatter lists of plots.
po4_scatter <- list()
for(i in unique(scatter_data$key)){
  po4_scatter[[paste0("PO4 X ", i)]] <- scatter_data %>%
  filter(key ==i) %>%
  ggplot(aes(log10(value), log10(`PO4 as P`), col = SUBSET), alpha = 0.05) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  scale_color_jama() + 
  labs(y = "log10(PO4)", x = paste0("log10(",i,")"), subtitle = paste0("PO4 X ", i))
  
}

ggarrange(plotlist = po4_scatter[c("PO4 X TN", "PO4 X TP" ,"PO4 X pH", "PO4 X DO", "PO4 X SP_COND")], ncol = 3, nrow = 2, common.legend = T, legend = "bottom")



```
---

```{r Fit an initial model, echo=F}
library(broom)
# Does PO4 explain resids of DOC ~ NO3 relationship.

ALL_CNP %>%
  ggplot(aes(log10(`NO3.as.N`), log10(DOC), col = ECO_TYPE)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  scale_color_jama() + 
  labs(y = "log10(DOC)", x = "log10(NO3)")
# Create model
docXno3 <- lm(log10(DOC) ~ log10(`NO3.as.N`) * ECO_TYPE, data = ALL_CNP)
# Summary
summary(docXno3)
# Extract resids and add to model.
ALL_CNP$mod_resids <- resid(docXno3)
# What is the effect of PO4?
ALL_CNP %>% ggplot(aes(log10(PO4.as.P), mod_resids, col = ECO_TYPE)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  scale_color_jama() + 
  labs(y = "Model residuals: log10(DOC) ~ log10(NO3)", x = "log10(PO4)") # Some funky stuff going on here...

# Create new model.
po4Xresid <- lm(mod_resids ~ log10(PO4.as.P) * ECO_TYPE, data = ALL_CNP)
summary(po4Xresid)

```

