---
title: "varycoef spatial regression"
author: "Daniel Gschwenter"
date: "4/15/2022"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE,  message = F, warning = F)

```

Code to generate spatial regression of NO~3~ ~ DOC + PO~4~ + DOC * ECOTYPE + PO~4~ * ECOTYPE. This is a preliminary model with low complexity. Trophic state will be added once it is generated as a variable for streams and lakes. 

The model will be fitted following the vignette for (varycoef)[https://cran.r-project.org/web/packages/varycoef/vignettes/example.html] where the model takes the form

$$ y = X \beta + W_n(s) + \epsilon $$

and W~n~(s) define the random effects that allowthe model coefficients to vary spatially.

```{r Setup}

lapply(c("tidyverse", "ggsci", "ggpubr", "sf", "tmap", "varycoef", "tmap"), require, character.only = T)

my_theme <- theme_bw() + 
  theme(axis.text = element_text(color = "black"), 
        axis.ticks = element_line(color = "black"), 
        panel.grid = element_blank(),
        strip.background = element_rect(size= 0.75, fill = "white"),
        legend.position = "bottom")

theme_set(my_theme)

```

```{r Load data}

all_cnp <- read_csv("https://raw.githubusercontent.com/swulfing/STOICH.Aim1/main/Data/ALL_CNP_April2022.csv")

summary(all_cnp)

```

```{r Remove low DOC sites for now}

# Log10 -1.5 is a convenient cutoff.
all_cnp %>% ggplot(aes(log10(DOC), log10(PO4.as.P))) + geom_point() 
all_cnp <- all_cnp %>% filter(log10(DOC) > -1.5)


```

```{r Create spatial object}

all_cnp <- st_as_sf(all_cnp, coords = c("LON", "LAT"), crs = 4326)

# Gonna reproject to projected CRS EPSG 26918; while this is centred over North America, this makes my life easier.


all_cnp <- st_transform(all_cnp, crs = 26918)

# Gonna export this and use the shapefile to compute spatial autocorrelation with python. Actually not sure this works with points so easily...
#st_write(all_cnp, "local_data/ALL_CNP.shp")

```

```{r Create map of objects}
# Data for world.
data("World")
st_crs(World)
World <- st_transform(World, 26918)
st_crs(World)
# Generate centroids.
#st_centroid(World) # Some errors.
World$valid <- st_is_valid(World)  # Check if geom is valid.
World %>% filter(valid != "TRUE") # Ok, 7 countries in Africa do not have valid geoms. We don't have data in these countries so we don't need to bother with them.
World <- World %>% filter(valid == "TRUE", continent %in% c("Europe", "North America"))
#World$centroid <- st_centroid(World)

```

```{r Map DOC}

tmap_leaflet(tm_shape(World) + tm_borders() + tm_shape(all_cnp) + tm_dots("DOC",  palette = "viridis", size = 0.1, breaks = c(5,10,20,30,40,60,80)))


```

```{r Map NO3}

tmap_leaflet(tm_shape(World) + tm_borders() + tm_shape(all_cnp) + tm_dots("NO3.as.N",  palette = "viridis", size = 0.1, breaks = c(0.1, 0.5, 1, 1.5, 2, 10, 30)))

```

```{r Map for PO4}

tmap_leaflet(tm_shape(World) + tm_borders() + tm_shape(all_cnp) + tm_dots("PO4.as.P",  palette = "viridis", size = 0.1, breaks = c(0.005, 0.01, 0.03, 0.05, 0.1, 0.5, 1, 4)))


```

---

Standard linear model (I want to see distribution of residuals)

```{r LM1}

lm_simple <- lm(log10(`NO3.as.N`) ~ log10(DOC) + log10(`PO4.as.P`) + ECO_TYPE *log10(DOC) + ECO_TYPE *log10(`PO4.as.P`), data = all_cnp)
lm_simple %>% summary()

plot(lm_simple, which = 1:2) # Some trends going on here.

```

```{r Spatial pattern of linear model residuals}

tmap_leaflet(tm_shape(World) + 
               tm_borders() + 
               tm_shape(
                 all_cnp %>% mutate(r = resid(lm_simple), r = (r - mean(r))/sd(r))) + # Standardises residuals...
               tm_dots("r", size = 0.1, breaks = seq(-10, 5, 1)))

# Quite a bit of clustering going on here...with some interesting outliers popping up here and there.
```


---

Fun and games are over, let's fit the spatial model.

```{r SVC model set up}
# Response.
y <- log10(all_cnp$NO3.as.N)
# Predictors
x <- model.matrix(~ 1 + log10(DOC) + log10(`PO4.as.P`) + ECO_TYPE *log10(DOC) + ECO_TYPE *log10(`PO4.as.P`), data = all_cnp )
# Locations
locs <- st_coordinates(all_cnp)/1000 # Converting to km (I  think?)
# Spatial weights: allowing effect of PO4 and DOC to vary spatially.
W <-  model.matrix(~ 1 + log10(DOC) + log10(`PO4.as.P`) + ECO_TYPE, data = all_cnp ) # Did not include interaction term here to make model more simple...

```


```{r Geostat model aka variogram}
library(gstat)
# Initial variogram
vgm1 <- variogram(log10(NO3.as.N) ~ 1, all_cnp)
plot(vgm1)
# Add estimates for what I think starting values should be.
vgm1 %>%
  ggplot(aes(dist, gamma)) + 
  geom_point() + 
  geom_smooth(se = F) +  # Loess smooth here  
  geom_vline(xintercept = 2e+06, col = "red") + 
  geom_hline(yintercept = 0.8, col = "red") + 
  geom_hline(yintercept = 0.3, col = "red")

```

```{r}
# Emprical variogram (based on residuals of linear model)
eV <- variogram(log10(NO3.as.N) ~ 1, data = all_cnp)
# Variogram model with initial values.
mV <- vgm(0.8, "Exp",2e+06 , 0.3)
# fit model 
(fV <- fit.variogram( eV, mV))
```


```{r Plot semivariogram}

plot(eV, model = fV) # Not too shabby I would say.

```

```{r Construct initial values for SVC model}

(init <- c(
  rep(c(
  # range
  fV$range[2], # Should already be in km,
  # variance
  fV$psill[2]),
  4),
  # nugget
  fV$psill[1]
))

```


```{r Control setting MLE}

control <- SVC_mle_control(
  profileLik = T, 
  init = init
)

```


```{r Fir SVC}

vc.fit <- SVC_mle(y = y, X = x, W = W, locs = locs, control = control) # Run without control for now; adding control = control throws this error: 
#Error in init_bounds_optim(control, p, q, id_obj, med_dist, y_var, OLS_mu) :  Initial values do not lie between lower and upper boundarys.

vc.fit %>% summary()

```

# Create subset of the data and fit SVC model.

```{r Random subset}

# Create random subset
rand <- all_cnp[sample(nrow(all_cnp), 1000), ]

table(rand$ECO_TYPE) # Ehhh not balanced but what the heck...

# Qucik map

tm_shape(World) + 
  tm_borders() +
tm_shape(rand) + 
  tm_dots()

```

```{r SVC model set up}
# Response.
y <- log10(rand$NO3.as.N)
# Predictors
x <- model.matrix(~ 1 + log10(DOC) + log10(`PO4.as.P`) + ECO_TYPE *log10(DOC) + ECO_TYPE *log10(`PO4.as.P`), data = rand )
# Locations
locs <- st_coordinates(rand)/1000 # Converting to km (I  think?)
# Spatial weights: allowing effect of PO4 and DOC to vary spatially.
W <-  model.matrix(~ 1 + log10(DOC) + log10(`PO4.as.P`) + ECO_TYPE, data = rand ) # Did not include interaction term here to make model more simple...

```


```{r Fir SVC for subset of data}

vc.fit <- SVC_mle(y = y, X = x, W = W, locs = locs)

vc.fit %>% summary()

```

```{r Coeffs}

coef(vc.fit)

```


For the SVC model with the random data subset following generalisations can be made:

1. NO~3~ decreases with DOC (significant at the 0.05 p level)

2. NO~3~ increases with PO~4~ (significant at the 0.05 p level)

3. Lakes generally have lower NO~3~/lower intercept for ~NO3~ (by a lot actually: about 0.6 mg/L)

4. Interaction of lake/river type with DOC is barely significant and increases NO~3~. 

5. Model captures a high degree of variance (R^2 = 0.79) but significance is low.

6. Interaction between PO~4~ and lake/river type was not significant.

7. Residuals are funky, especially at tail ends of distributions.

8. Large range in spatially varying coeffs....

  1. How useful are these spatially variable coefs if range is so big?
  
  2. Can we draw meaningful insights if range is too large?
  
  3. Are we interested in prediction or inference? If inference, do we need coefs to vary spatially? We are more interested in the behaviour of the systems in general than prediction NO~3~ at a given location (are we not?)...
  
9. Random model with subset of data and no control took about an hour to fit....I used 1,000 observations and a relatively simple model.... data set has >16,000 observations fyi...

10. Lack of sign. for most of the terms suggests this model isn't top notch/our approach may not be suitable for the data/things are getting messy...

```{r Check residuals}

plot(vc.fit, which = 1:2)

```

```{r Extract residuals}

rand$model_resids <- residuals(vc.fit)

tmap_mode("view")
tm_shape(World) + 
  tm_borders() + 
tm_shape(rand) + 
  tm_dots(col = "model_resids",shape = "ECO_TYPE",  palette = "Spectral", size = 0.1)

```


```{r Playing with resids}


rand %>% 
  ggplot(aes(ECO_TYPE, model_resids)) + 
  geom_boxplot()

rand %>% 
  ggplot(aes(PO4.as.P, model_resids, col = ECO_TYPE)) + 
  geom_point()


rand %>% 
  ggplot(aes(DOC, model_resids, col = ECO_TYPE)) + 
  geom_point()

```

