---
title: "Quantile_Categories"
date: "January 5, 2022"
output:
  html_document:
    df_print: paged
---

```{r setup, inlcude = FALSE, warning = FALSE}

library(tidyverse)
library(MetBrewer)
library(sf)
library(ggspatial)
library(raster)
library(rnaturalearth)
library(rgeos)

```

## Preparing the data, looking at a boxplot

```{r, warning = F, message = F, fig.width = 6.25, fig.height = 4.25}
setwd("C:/Users/linne/OneDrive - University of Wyoming/PhD_code/STOICH.Aim1")
source("Code/masterData.R")

medians <- ALL_CNP %>%
  group_by(SITE_ID, LAT, LON, ECO_TYPE) %>%
  summarise(m.DOC = median(DOC),
            m.NO3 = median(NO3.as.N),
            m.TP = median(TP))


##boxplot of data####
medians.box <- medians %>%
  rename(DOC = m.DOC,
         NO3 = m.NO3,
         TP = m.TP) %>%
  pivot_longer(5:7, names_to = "Analyte", values_to = "Median")
  

ggplot(medians.box) +
  geom_boxplot(aes(Analyte, Median)) +
  scale_y_log10()

```

## Checking out the quantiles

```{r, warning = F, message = F, fig.width = 6.25, fig.height = 4.25}
QN <- summary(medians$m.NO3)
QP <- summary(medians$m.TP)
QC <- summary(medians$m.DOC)

QN
QP
QC

```

## Creating categories based on quantiles

*very low == < 1st quantile 
*low == between 1st and 3rd quantile
*medium == between 3rd and 4th quantile
*very high == greater than 4th quantile

```{r, warning = F, message = F, fig.width = 6.25, fig.height = 4.25}
CNP_categories <- medians %>%
  mutate(NO3_limit = NA,
         TP_limit = NA,
         DOC_limit = NA) %>%
  mutate(NO3_limit = ifelse(m.NO3 < QN[2], "very low", 
                            ifelse(between(m.NO3, QN[2], QN[3]), "low",
                                   ifelse(between(m.NO3, QN[3], QN[4]), "medium", 
                                   ifelse(m.NO3 > QN[4], "high", NO3_limit))))) %>%
  mutate(TP_limit = ifelse(m.TP < QP[2], "very low", 
                            ifelse(between(m.TP, QP[2], QP[3]), "low",
                                   ifelse(between(m.TP, QP[3], QP[4]), "medium", 
                                          ifelse(m.TP > QP[4], "high", TP_limit)))))%>%
  mutate(DOC_limit = ifelse(m.DOC < QC[2], "very low", 
                           ifelse(between(m.DOC, QC[2], QC[3]), "low",
                                  ifelse(between(m.DOC, QC[3], QC[4]), "medium", 
                                         ifelse(m.DOC > QC[4], "high", DOC_limit)))))

#create factors so that thresholds are put into an order that makes sense on figures
CNP_categories$NO3_limit <- factor(CNP_categories$NO3_limit,
                                      levels = c("high", "medium", "low", "very low"))
CNP_categories$TP_limit <- factor(CNP_categories$TP_limit,
                                   levels = c("high", "medium", "low", "very low"))
CNP_categories$DOC_limit <- factor(CNP_categories$DOC_limit,
                                   levels = c("high", "medium", "low", "very low"))


```

## Plotting relationships

#### Full set of data

```{r, warning = F, message = F,  fig.width = 6.25, fig.height = 4.25}

ggplot(CNP_categories) +
  geom_point(aes(m.DOC, m.NO3, color = TP_limit)) +
  scale_color_manual(values = met.brewer("Egypt"), "TP Limits") +
  geom_hline(yintercept = QN[2]) + #adding lines to the plot for NO3, DOC "limits"
  geom_hline(yintercept = QN[3]) +
  geom_hline(yintercept = QN[4]) +
  geom_vline(xintercept = QC[2]) +
  geom_vline(xintercept = QC[3]) +
  geom_vline(xintercept = QC[4]) +
  theme_bw() +
  labs(x = "DOC"~mg~L^-1,
       y = "Nitrate (N)"~mg~L^-1)

```

#### "Zooming in" on the above figure

```{r, warning = F, message = F,  fig.width = 6.25, fig.height = 4.25}

ggplot(CNP_categories) +
  geom_point(aes(m.DOC, m.NO3, color = TP_limit)) +
  scale_color_manual(values = met.brewer("Egypt"), "TP Limits") +
  geom_hline(yintercept = QN[2]) + #adding lines to the plot for NO3, DOC "limits"
  geom_hline(yintercept = QN[3]) +
  geom_hline(yintercept = QN[4]) +
  geom_vline(xintercept = QC[2]) +
  geom_vline(xintercept = QC[3]) +
  geom_vline(xintercept = QC[4]) +
  theme_bw() +
  labs(x = "DOC"~mg~L^-1,
       y = "Nitrate (N)"~mg~L^-1) +
  xlim(0,30) + ylim(0,10)

```


#### NO3-TP, showing the nitrate thresholds

```{r, warning = F, message = F, fig.width = 6.25, fig.height = 4.25}

ggplot(CNP_categories) +
  geom_point(aes(m.TP, m.NO3, color = NO3_limit)) +
  scale_color_manual(values = met.brewer("Egypt"), "NO3 Limits") +
  theme_bw() +
  labs(x = "TP"~mg~L^-1,
       y = "Nitrate (N)"~mg~L^-1)

```


#### NO3-TP, showing the phosphorus thresholds

```{r, warning = F, message = F,  fig.width = 6.25, fig.height = 4.25}

ggplot(CNP_categories) +
  geom_point(aes(m.TP, m.NO3, color = TP_limit)) +
  scale_color_manual(values = met.brewer("Egypt"), "TP Limits") +
  theme_bw() +
  labs(x = "TP"~mg~L^-1,
       y = "Nitrate (N)"~mg~L^-1)

```

## Mapping -- any spatial observations?

```{r, warning = F, message = F,  fig.width = 6.25, fig.height = 4.25}
world <- ne_countries(returnclass = "sf")

#TP data
tp <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(CNP_categories, mapping = aes(LON, LAT, color = TP_limit)) +
  scale_color_manual(values = met.brewer("Egypt")) +
  theme_bw() +
  labs(x = "Longitude",
       y = "Latitude",
       title = "TP") +
  theme(legend.title = element_blank()) +
  coord_sf(xlim = c(-171.565140, 49.021941), ylim = c(10.942356, 88.306487), expand = FALSE)


#NO3 data
no3 <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(CNP_categories, mapping = aes(LON, LAT, color = NO3_limit)) +
  scale_color_manual(values = met.brewer("Egypt")) +
  theme_bw() +
  labs(x = "Longitude",
       y = "Latitude",
       title = "NO3") +
  theme(legend.title = element_blank()) +
  coord_sf(xlim = c(-171.565140, 49.021941), ylim = c(10.942356, 88.306487), expand = FALSE)


#DOC data
doc <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(CNP_categories, mapping = aes(LON, LAT, color = DOC_limit)) +
  scale_color_manual(values = met.brewer("Egypt")) +
  theme_bw() +
  labs(x = "Longitude",
       y = "Latitude",
       title = "DOC") +
  theme(legend.title = element_blank()) +
  coord_sf(xlim = c(-171.565140, 49.021941), ylim = c(10.942356, 88.306487), expand = FALSE)

tp
no3
doc

```
