---
title: "Trophic States"
author: "Linnea"
date: '2022-04-21'
output: html_document
---

```{r setup, include=FALSE, message = FALSE, echo = FALSE}
library(sf)
library(ggspatial)
library(raster)
library(rnaturalearth)
library(rgeos)
library(patchwork)
```


```{r, warning = FALSE, echo = FALSE, inlcude = FALSE, message = FALSE}
source("C:/Users/lrock1/Downloads/PhD_code/STOICH.Aim1/Code/masterDataVARIABLES.R")
```

## How many observations of concurrently collected C, N, P with a trophic state associated?
```{r, warning = FALSE, message = FALSE}
total.TS <- ALL_CNP |>
  count(TROPHIC_STATE)

total.TS
```
We had a total of 106920 observations with a TP measurement to calculate trophic state. 6342 sites/dates did not have a TP measurement. 

The trophic state was calculated as follows: 

 -- OLIGOTROPHIC (TP <= 10 ug/L)
 -- MESOTROPHIC (10 > TP >= 25 ug/L)
 -- EUTROPHIC (TP > 25 ug/L)
 
## Maps showing trophic states of lakes and rivers across our regions
```{r, warning = FALSE, include = FALSE}
world <- ne_countries(returnclass = "sf")

america.map <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON < -50), mapping = aes(LON, LAT, color = TROPHIC_STATE)) +
  theme_light() +
  scale_color_viridis_d() +
  coord_sf(xlim = c(-171.565140, -50), ylim = c(10.942356, 80), expand = FALSE)
  


europe.map <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON > -50), mapping = aes(LON, LAT, color = TROPHIC_STATE)) +
  theme_light() +
  scale_color_viridis_d() +
  coord_sf(xlim = c(-40, 49.021941), ylim = c(30, 70), expand = FALSE)

```

```{r, warning = FALSE, message = FALSE}
america.map 
europe.map
```

## Eutrophic waters only - PO4
Note that I am taking natural log of PO4
```{r message=FALSE, warning=FALSE, include = FALSE}

america.map.P <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON < -50 & TROPHIC_STATE == "Eutrophic"), mapping = aes(LON, LAT, color = log(PO4.as.P))) +
  theme_light() +
  scale_color_viridis_c() +
  coord_sf(xlim = c(-171.565140, -50), ylim = c(10.942356, 80), expand = FALSE)
  


europe.map.P <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON > -50 & TROPHIC_STATE == "Eutrophic"), mapping = aes(LON, LAT, color = log(PO4.as.P))) +
  theme_light() +
  scale_color_viridis_c() +
  coord_sf(xlim = c(-40, 49.021941), ylim = c(30, 70), expand = FALSE)
```

```{r, warning = FALSE, message = FALSE}
america.map.P 
europe.map.P
```

## Mesotrophic waters only - PO4
Note that I am taking natural log of PO4
```{r, warning = FALSE, message = FALSE, include = FALSE}

america.map.P <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON < -50 & TROPHIC_STATE == "Mesotrophic"), mapping = aes(LON, LAT, color = log(PO4.as.P))) +
  theme_light() +
  scale_color_viridis_c() +
  coord_sf(xlim = c(-171.565140, -50), ylim = c(10.942356, 80), expand = FALSE)
  

europe.map.P <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON > -50 & TROPHIC_STATE == "Mesotrophic"), mapping = aes(LON, LAT, color = log(PO4.as.P))) +
  theme_light() +
  scale_color_viridis_c() +
  coord_sf(xlim = c(-40, 49.021941), ylim = c(30, 70), expand = FALSE)
```

```{r, warning = FALSE, message = FALSE}
america.map.P 
europe.map.P
```

## Oligotrophic waters only - PO4
Note that I am taking natural log of PO4
```{r, warning = FALSE, message = FALSE, include=FALSE}

america.map.P <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON < -50 & TROPHIC_STATE == "Oligotrohpic"), mapping = aes(LON, LAT, color = log(PO4.as.P))) +
  theme_light() +
  scale_color_viridis_c() +
  coord_sf(xlim = c(-171.565140, -50), ylim = c(10.942356, 80), expand = FALSE)
  

europe.map.P <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON > -50 & TROPHIC_STATE == "Oligotrophic"), mapping = aes(LON, LAT, color = log(PO4.as.P))) +
  theme_light() +
  scale_color_viridis_c() +
  coord_sf(xlim = c(-40, 49.021941), ylim = c(30, 70), expand = FALSE)
```

```{r, warning = FALSE, message = FALSE}
america.map.P 
europe.map.P
```


## Eutrophic waters only - NO3
Note that I am taking natural log of NO3
```{r, warning = FALSE, message = FALSE, inlcude = FALSE}

america.map.N <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON < -50 & TROPHIC_STATE == "Eutrophic"), mapping = aes(LON, LAT, color = log(NO3.as.N))) +
  theme_light() +
  scale_color_viridis_c() +
  coord_sf(xlim = c(-171.565140, -50), ylim = c(10.942356, 80), expand = FALSE)
  


europe.map.N <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON > -50 & TROPHIC_STATE == "Eutrophic"), mapping = aes(LON, LAT, color = log(NO3.as.N))) +
  theme_light() +
  scale_color_viridis_c() +
  coord_sf(xlim = c(-40, 49.021941), ylim = c(30, 70), expand = FALSE)
```

```{r, warning = FALSE, message = FALSE}
america.map.N 
europe.map.N
```

## Mesotrophic waters only - NO3
Note that I am taking natural log of NO3
```{r, warning = FALSE, message = FALSE, include=FALSE}

america.map.N <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON < -50 & TROPHIC_STATE == "Mesotrophic"), mapping = aes(LON, LAT, color = log(NO3.as.N))) +
  theme_light() +
  scale_color_viridis_c() +
  coord_sf(xlim = c(-171.565140, -50), ylim = c(10.942356, 80), expand = FALSE)
  
europe.map.N <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON > -50 & TROPHIC_STATE == "Mesotrophic"), mapping = aes(LON, LAT, color = log(NO3.as.N))) +
  theme_light() +
  scale_color_viridis_c() +
  coord_sf(xlim = c(-40, 49.021941), ylim = c(30, 70), expand = FALSE)
```

```{r, warning = FALSE, message = FALSE}
america.map.N 
europe.map.N
```

## Oligotrophic waters only - NO3
Note that I am taking natural log of NO3
```{r, warning = FALSE, message = FALSE, inlcude = FALSE}

america.map.N <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON < -50 & TROPHIC_STATE == "Oligotrohpic"), mapping = aes(LON, LAT, color = log(NO3.as.N))) +
  theme_light() +
  scale_color_viridis_c() +
  coord_sf(xlim = c(-171.565140, -50), ylim = c(10.942356, 80), expand = FALSE)
  


europe.map.N <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON > -50 & TROPHIC_STATE == "Oligotrophic"), mapping = aes(LON, LAT, color = log(NO3.as.N))) +
  theme_light() +
  scale_color_viridis_c() +
  coord_sf(xlim = c(-40, 49.021941), ylim = c(30, 70), expand = FALSE)
```

```{r, warning = FALSE, message = FALSE}
america.map.N 
europe.map.N
```



## Eutrophic waters only - DOC
Note that I am taking natural log of DOC
```{r, warning = FALSE, message = FALSE, include=FALSE}

america.map.C <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON < -50 & TROPHIC_STATE == "Eutrophic"), mapping = aes(LON, LAT, color = log(DOC))) +
  theme_light() +
  scale_color_viridis_c() +
  coord_sf(xlim = c(-171.565140, -50), ylim = c(10.942356, 80), expand = FALSE)
  
europe.map.C <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON > -50 & TROPHIC_STATE == "Eutrophic"), mapping = aes(LON, LAT, color = log(DOC))) +
  theme_light() +
  scale_color_viridis_c() +
  coord_sf(xlim = c(-40, 49.021941), ylim = c(30, 70), expand = FALSE)
```

```{r, warning = FALSE, message = FALSE}
america.map.C 
europe.map.C
```

## Mesotrophic waters only - DOC
Note that I am taking natural log of DOC
```{r, warning = FALSE,message=FALSE, include=FALSE}

america.map.C <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON < -50 & TROPHIC_STATE == "Mesotrophic"), mapping = aes(LON, LAT, color = log(DOC))) +
  theme_light() +
  scale_color_viridis_c() +
  coord_sf(xlim = c(-171.565140, -50), ylim = c(10.942356, 80), expand = FALSE)
  


europe.map.C <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON > -50 & TROPHIC_STATE == "Mesotrophic"), mapping = aes(LON, LAT, color = log(DOC))) +
  theme_light() +
  scale_color_viridis_c() +
  coord_sf(xlim = c(-40, 49.021941), ylim = c(30, 70), expand = FALSE)
```

```{r, warning = FALSE, message = FALSE}
america.map.C 
europe.map.C
```

## Oligotrophic waters only - DOC
Note that I am taking natural log of DOC
```{r, warning = FALSE, message=FALSE, include=FALSE}

america.map.C <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON < -50 & TROPHIC_STATE == "Oligotrohpic"), mapping = aes(LON, LAT, color = log(DOC))) +
  theme_light() +
  scale_color_viridis_c() +
  coord_sf(xlim = c(-171.565140, -50), ylim = c(10.942356, 80), expand = FALSE)
  


europe.map.C <- ggplot() +
  geom_sf(world, mapping = aes(), fill = "white") +
  geom_point(ALL_CNP |> filter(LON > -50 & TROPHIC_STATE == "Oligotrophic"), mapping = aes(LON, LAT, color = log(DOC))) +
  theme_light() +
  scale_color_viridis_c() +
  coord_sf(xlim = c(-40, 49.021941), ylim = c(30, 70), expand = FALSE)
```

```{r, warning = FALSE, message = FALSE}
america.map.C 
europe.map.C
```
