---
title: "Fit GAM to cnp data"
author: "David Nguyen"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# gams for manuscript
library(tidyverse)
library(mgcv)

```

# Overview

I fit GAMs of the form

\begin{align*}
\log_{10}(y_i) &= \beta_0 + \beta_1 \text{ECOTYPE}\\ 
               &+ f_1(\log_{10}DOC_i) \times \text{ECOTYPE} + f_2(\log_{10}PO_4) \times \text{ECOTYPE}\\ 
               &+ f_3(\log_{10}DOC \times \log_{10}PO_4)  \epsilon_i
\end{align*}

<!-- $$\log_{10}(y_i) = \beta_0 + \beta_1 \text{ECOTYPE} + f_1(\log_{10}DOC_i) \times \text{ECOTYPE} + f_2(\log_{10}PO_4) \times \text{ECOTYPE} + f_3(\log_{10}DOC \times \log_{10}PO_4)  \epsilon_i$$ -->

Where $y_i$ is the annual median NO3 for site $i$. The functions $f(.)$ are smooth functions. ECOTYPE is an indicator which determines whether the site is a lake or a river/stream. The error is assumed iid normal (on the log-scale) with zero mean.

Using the `mgcv` package I fit this model using the syntax: `gam(y ~ s(DOC, by = ECO_TYPE) + s(PO4, by = ECO_TYPE) + ti(DOC, PO4) + ECO_TYPE)`

The `mgcv::ti()` function can be used to specify the r formula equivalent of the interaction `DOC:PO4`.

I suppose I should include ECO_TYPE along with the interaction term.

```{r}
# read in data
cnp <- read.csv("data/ALL_CNP_med.csv")
cnp$ECO_TYPE <- factor(cnp$ECO_TYPE) # for fitting gam by s(.., by = ECO_TYPE)

# merge median of medians of PO4 (rivers) and TP (lakes) into one variable
# this is so we can fit a common model for lakes and rivers
cnp <- cnp %>% mutate(phos_med = ifelse(ECO_TYPE == "River/Stream", 
                                                 P_med,
                                                 TP_med))
# appropriate P measurment missing for 5 EU lakes
# cnp_med %>% group_by(ECO_TYPE, region) %>% summarise(missing_phos = sum(is.na(phos_med)))
cnp <- cnp %>% filter(!is.na(phos_med))

# log everything
cnp <- cnp %>% mutate(log_N_med = log(N_med),
                              log_phos_med = log(phos_med),
                              log_DOC_med = log(DOC_med))

# define Europe vs NA subset
cnp$region <- factor(ifelse(cnp$LON < -50, "North America", "Europe"))

# split into NA and EU subset
cnp_med_na <- filter(cnp, region == "North America")
cnp_med_eu <- filter(cnp, region == "Europe")

### CHeck that subsetting was correst
# world_map <- map_data("world")
# cnp %>%
# ggplot() +
#   geom_map(aes(map_id = region), col = "gray", fill = "white", 
#            data = world_map, map = world_map) +
#   geom_point(aes(
#     x = LON,
#     y = LAT,
#     col = region))

gam_eu <- readRDS("models/gam_eu.RDS")
gam_na <- readRDS("models/gam_na.RDS")
```


```{r}
cnp %>% group_by(region, ECO_TYPE) %>% summarise(nsites = n()) %>%
  knitr::kable(caption = "Number of sites per region and type")
```

# Europe

```{r}
# EU region
# fit 2d smooth or 2 separate smooths
# both have interaction with ECO_TYPE and main effect
# gam_eu <- gam(log_N_med ~ s(log_DOC_med, by = ECO_TYPE, bs = "fs") + 
#                             s(log_phos_med, by = ECO_TYPE, bs = "fs", k = 30) + 
#                             ti(log_DOC_med, log_phos_med) + 
#                             ECO_TYPE,
#     data = filter(cnp, region == "Europe"), method = "REML") 
# saveRDS(gam_eu, file = "models/gam_eu.RDS")


### gam with spatially varying intercept
gam_eu_spatial <- gam(log_N_med ~ s(log_DOC_med, by = ECO_TYPE, bs = "fs") +
                            s(log_phos_med, by = ECO_TYPE, bs = "fs", k = 30) +
                            ti(log_DOC_med, log_phos_med) +
                            ECO_TYPE +
                s(LON, LAT, k = 100),
    data = filter(cnp, region == "Europe"), method = "REML")

# saveRDS(gam_eu_spatial, file = "models/gam_eu_spatial.RDS")
gam_eu_spatial <- readRDS("models/gam_eu_spatial.RDS")

summary(gam_eu_spatial)

summary(gam_eu)

plot(gam_eu_spatial, pages = 1, scheme = 2)
plot(gam_eu, pages = 1, scheme = 2)

gam.check(gam_eu_spatial)
```


```{r}
### gam with spatially varying intercept
# gam_na_spatial <- gam(log_N_med ~ s(log_DOC_med, by = ECO_TYPE, bs = "fs") +
#                             s(log_phos_med, by = ECO_TYPE, bs = "fs", k = 30) +
#                             ti(log_DOC_med, log_phos_med) +
#                             ECO_TYPE +
#                 s(LON, LAT, k = 100),
#     data = filter(cnp, region == "North America"), method = "REML")
# 
# saveRDS(gam_na_spatial, file = "models/gam_na_spatial.RDS")
gam_na_spatial <- readRDS("models/gam_na_spatial.RDS")

summary(gam_na_spatial)

summary(gam_na)

plot(gam_na_spatial, pages = 1, scheme = 2)
plot(gam_na, pages = 1, scheme = 2)

gam.check(gam_na_spatial)
```


```{r}
world_map <- map_data("world")

cnp_med_na$gam_resid <- residuals(gam_na)
cnp_med_eu$gam_resid <- residuals(gam_eu)

cnp_med_na$gam_resid_sp <- residuals(gam_na_spatial)
cnp_med_eu$gam_resid_sp <- residuals(gam_eu_spatial)

cnp_med_na <- cnp_med_na %>% mutate(sign_resid = ifelse(gam_resid < 0, "-", "+"),
                                    sign_resid_sp = ifelse(gam_resid_sp < 0, "-", "+"))
cnp_med_eu <- cnp_med_eu %>% mutate(sign_resid = ifelse(gam_resid < 0, "-", "+"),
                                    sign_resid_sp = ifelse(gam_resid_sp < 0, "-", "+"))

pr_na <- cnp_med_na %>% 
  ggplot() +
  geom_map(aes(map_id = region), col = "gray", fill = "white",
           data = world_map, map = world_map) +
  geom_point(aes(
    x = LON,
    y = LAT,
    col = sign_resid,
    size = abs(gam_resid)),
    alpha = 0.05, shape = 16) +
  viridis::scale_color_viridis(discrete = T) +
  facet_wrap(~ECO_TYPE) +
  labs(title = "Residual",
       col = "Sign")+ 
  guides(colour = guide_legend(override.aes = list(alpha = 1)),
                              size = guide_legend(override.aes = list(alpha = 1)))


pr_eu <- cnp_med_eu %>% 
  ggplot() +
  geom_map(aes(map_id = region), col = "gray", fill = "white",
           data = world_map, map = world_map) +
  geom_point(aes(
    x = LON,
    y = LAT,
    col = sign_resid,
    size = abs(gam_resid)),
    alpha = 0.05, shape = 16) +
  viridis::scale_color_viridis(discrete = T) +
  facet_wrap(~ECO_TYPE) +
  labs(title = "Residual",
       col = "Sign") + 
  guides(colour = guide_legend(override.aes = list(alpha = 1)),
                              size = guide_legend(override.aes = list(alpha = 1)))

pr_na_sp <- cnp_med_na %>% 
  ggplot() +
  geom_map(aes(map_id = region), col = "gray", fill = "white",
           data = world_map, map = world_map) +
  geom_point(aes(
    x = LON,
    y = LAT,
    col = sign_resid_sp,
    size = abs(gam_resid_sp)),
    alpha = 0.05, shape = 16) +
  viridis::scale_color_viridis(discrete = T) +
  facet_wrap(~ECO_TYPE) +
  labs(title = "Residual",
       col = "Sign")+ 
  guides(colour = guide_legend(override.aes = list(alpha = 1)),
                              size = guide_legend(override.aes = list(alpha = 1)))


pr_eu_sp <- cnp_med_eu %>% 
  ggplot() +
  geom_map(aes(map_id = region), col = "gray", fill = "white",
           data = world_map, map = world_map) +
  geom_point(aes(
    x = LON,
    y = LAT,
    col = sign_resid_sp,
    size = abs(gam_resid_sp)),
    alpha = 0.05, shape = 16) +
  viridis::scale_color_viridis(discrete = T) +
  facet_wrap(~ECO_TYPE) +
  labs(title = "Residual",
       col = "Sign") + 
  guides(colour = guide_legend(override.aes = list(alpha = 1)),
                              size = guide_legend(override.aes = list(alpha = 1)))

library(patchwork)
pr_na + pr_na_sp + pr_eu + pr_eu_sp
```

### Predicted NO3 maps

#### North America

We can easily see in the map that predicted NO[3] median of medians is lower in lakes relative to rivers.

```{r}
cnp_med_na %>% 
  ggplot() +
  geom_map(aes(map_id = region), col = "gray", fill = "white",
           data = world_map, map = world_map) +
  geom_point(aes(
    x = LON,
    y = LAT,
    col = lm_fit),
    alpha = 0.5, shape = 1) +
  scale_color_viridis() +
  facet_wrap(~ECO_TYPE) +
  labs(title = "Predicted log(NO3)",
       col = "Prediction")


### Compare DOC x PO4 interaction model with main effects model
```

```{r eval = FALSE}
# interaction model better by AIC -->
AIC(gam_eu_interaction, gam_eu_main) -->
``` 

### Plot regression predictions

```{r}
quant_log_phos_na <- quantile(cnp_med_na$log_phos_med, c(0.25, 0.5, 0.75))
quant_log_phos_eu <- quantile(cnp_med_eu$log_phos_med, c(0.25, 0.5, 0.75))
         
nd_DOC_na <- expand_grid(ECO_TYPE = c("River/Stream", "Lake"),
                  # `log(DOC_med)` = seq(from = log(min(cnp_med_na$DOC_med)), 
                  #                   to = log(max(cnp_med_na$DOC_med)),
                  #                   length = 100),
                  # `log(phos_med)` = log(quantile(cnp_med_na$phos_med, 
                  #                                c(0.25, 0.5, 0.75)))
                  log_DOC_med = seq(from = min(cnp_med_na$log_DOC_med), 
                                    to = max(cnp_med_na$log_DOC_med),
                                    length = 100),
                  log_phos_med = quant_log_phos_na
                  )
preds_DOC_na <- predict(gam_na, newdata = nd_DOC_na, se.fit = TRUE)
nd_DOC_na$pred <- preds_DOC_na$fit
nd_DOC_na$se <- preds_DOC_na$se.fit
nd_DOC_na <- nd_DOC_na %>% mutate(conf.low = pred - 2 * se,
                     conf.high = pred + 2 * se)

nd_DOC_eu <- expand_grid(ECO_TYPE = c("River/Stream", "Lake"),
                  # `log(DOC_med)` = seq(from = log(min(cnp_med_eu$DOC_med)), 
                  #                   to = log(max(cnp_med_eu$DOC_med)),
                  #                   length = 100),
                  # `log(phos_med)` = log(quantile(cnp_med_eu$phos_med, 
                  #                                c(0.25, 0.5, 0.75)))
                  log_DOC_med = seq(from = min(cnp_med_eu$log_DOC_med), 
                                    to = max(cnp_med_eu$log_DOC_med),
                                    length = 100),
                  log_phos_med = quant_log_phos_eu
                  )
preds_DOC_eu <- predict(gam_eu, newdata = nd_DOC_eu, se.fit = TRUE)
nd_DOC_eu$pred <- preds_DOC_eu$fit
nd_DOC_eu$se <- preds_DOC_eu$se.fit
nd_DOC_eu <- nd_DOC_eu %>% mutate(conf.low = pred - 2 * se,
                     conf.high = pred + 2 * se)

nd_DOC_eu <- nd_DOC_eu %>% mutate(region = "Europe",
                                  p_quantile = case_when(
                                    log_phos_med == quant_log_phos_eu[1] ~ "25%",
                                    log_phos_med == quant_log_phos_eu[2] ~ "50%",
                                    log_phos_med == quant_log_phos_eu[3] ~ "75%",
                                                         ))
nd_DOC_na <- nd_DOC_na %>% mutate(region = "North America",
                                  p_quantile = case_when(
                                    log_phos_med == quant_log_phos_na[1] ~ "25%",
                                    log_phos_med == quant_log_phos_na[2] ~ "50%",
                                    log_phos_med == quant_log_phos_na[3] ~ "75%",
                                                         ))
nd_DOC_all <- rbind(nd_DOC_eu, nd_DOC_na)

nd_DOC_all$model <- "GAM"
# write_csv(nd_DOC_all, "pred_DOC_gam")
```

```{r}
nd_DOC_all %>%
  ggplot(aes(x = log_DOC_med, y = pred)) +
  geom_line(aes(linetype = ECO_TYPE), size = 1.5) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = ECO_TYPE), alpha = 0.5) +
  viridis::scale_fill_viridis(discrete = T) +
  facet_grid(p_quantile~region)
```

```{r}
quant_log_DOC_na <- quantile(cnp_med_na$log_DOC_med, c(0.25, 0.5, 0.75))
quant_log_DOC_eu <- quantile(cnp_med_eu$log_DOC_med, c(0.25, 0.5, 0.75))
         
nd_phos_na <- expand_grid(ECO_TYPE = c("River/Stream", "Lake"),
                  # `log(phos_med)` = seq(from = log(min(cnp_med_na$phos_med)), 
                  #                   to = log(max(cnp_med_na$phos_med)),
                  #                   length = 100),
                  # `log(phos_med)` = log(quantile(cnp_med_na$phos_med, 
                  #                                c(0.25, 0.5, 0.75)))
                  log_phos_med = seq(from = min(cnp_med_na$log_phos_med), 
                                    to = max(cnp_med_na$log_phos_med),
                                    length = 100),
                  log_DOC_med = quant_log_DOC_na
                  )
preds_phos_na <- predict(gam_na, newdata = nd_phos_na, se.fit = TRUE)
nd_phos_na$pred <- preds_phos_na$fit
nd_phos_na$se <- preds_phos_na$se.fit
nd_phos_na <- nd_phos_na %>% mutate(conf.low = pred - 2 * se,
                     conf.high = pred + 2 * se)

nd_phos_eu <- expand_grid(ECO_TYPE = c("River/Stream", "Lake"),
                  # `log(phos_med)` = seq(from = log(min(cnp_med_eu$phos_med)), 
                  #                   to = log(max(cnp_med_eu$phos_med)),
                  #                   length = 100),
                  # `log(phos_med)` = log(quantile(cnp_med_eu$phos_med, 
                  #                                c(0.25, 0.5, 0.75)))
                  log_phos_med = seq(from = min(cnp_med_eu$log_phos_med), 
                                    to = max(cnp_med_eu$log_phos_med),
                                    length = 100),
                  log_DOC_med = quant_log_DOC_eu
                  )
preds_phos_eu <- predict(gam_eu, newdata = nd_phos_eu, se.fit = TRUE)
nd_phos_eu$pred <- preds_phos_eu$fit
nd_phos_eu$se <- preds_phos_eu$se.fit
nd_phos_eu <- nd_phos_eu %>% mutate(conf.low = pred - 2 * se,
                     conf.high = pred + 2 * se)

nd_phos_eu <- nd_phos_eu %>% mutate(region = "Europe",
                                  DOC_quantile = case_when(
                                    log_DOC_med == quant_log_DOC_eu[1] ~ "25%",
                                    log_DOC_med == quant_log_DOC_eu[2] ~ "50%",
                                    log_DOC_med == quant_log_DOC_eu[3] ~ "75%",
                                                         ))
nd_phos_na <- nd_phos_na %>% mutate(region = "North America",
                                  DOC_quantile = case_when(
                                    log_DOC_med == quant_log_DOC_na[1] ~ "25%",
                                    log_DOC_med == quant_log_DOC_na[2] ~ "50%",
                                    log_DOC_med == quant_log_DOC_na[3] ~ "75%",
                                                         ))

nd_phos_all <- rbind(nd_phos_eu, nd_phos_na)

nd_phos_all$model <- "GAM"
# write_csv(nd_phos_all, "pred_phos_gam.csv")
```

```{r}
nd_phos_all %>%
  ggplot(aes(x = log_phos_med, y = pred)) +
  geom_line(aes(linetype = ECO_TYPE), size = 1.5) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = ECO_TYPE), alpha = 0.5) +
  viridis::scale_fill_viridis(discrete = T) +
  facet_grid(DOC_quantile~region)
```


### Plot smooths

<!-- Note that all "smooth" plots are partial effect plots, e.g., they are centered on zero. Not the same as prediction plots. -->

```{r}
# how to make better plots?
plot(gam_eu, scheme = 2#
     ,page = 1
     ,shift = coef(gam_eu)[1]
     )
```

### Plot predicted average median log(NO3)

First level of ECO_TYpe is "lake" and second level is "River/Stream.

Pay attention to direction of axis when interpreting (should extract all this and make better plots instead). 

```{r}
vis.gam(gam_eu, view = c("ECO_TYPE", "log_DOC_med"), theta = 240)
vis.gam(gam_eu, view = c("ECO_TYPE", "log_phos_med"), theta = 60)
```

### Model checking

Model assumptions are reasonably well met.

May have slight heteroskedasticity, low predicted values have smaller error than high predicted values (see residual vs fitted plot).

Values of basis dimension k are ok since edf < k.

```{r}
# Seems fine wrt to basis functions. edf < k for all smooths
gam.check(gam_eu)
```

# North America


```{r}
# NA region
# fit 2d smooth or 2 separate smooths
# both have interaction with ECO_TYPE and main effect
# gam_NA_interaction <- gam(log10(N_med) ~ s(log10(DOC_med), by = ECO_TYPE) + s(log10(P_med), by = ECO_TYPE) + ti(log10(DOC_med), log10(P_med)) + ECO_TYPE,
#                           data = filter(cnp, region == "North America"), method = "ML") 
# 
# gam_NA_main <- gam(log10(N_med) ~ s(log10(DOC_med), by = ECO_TYPE) + s(log10(P_med), by = ECO_TYPE) + ECO_TYPE,
#                    data = filter(cnp, region == "North America"), method = "ML") 
# 
# summary(gam_NA_interaction)
# summary(gam_NA_main)

# gam_na <- gam(log_N_med ~ s(log_DOC_med, by = ECO_TYPE, bs = "fs") + 
#                             s(log_phos_med, by = ECO_TYPE, bs = "fs", k = 30) + 
#                             ti(log_DOC_med, log_phos_med) + 
#                             ECO_TYPE,
#                    data = filter(cnp, region == "North America"), method = "REML") 
# saveRDS(gam_na, file = "models/gam_na.RDS")
summary(gam_na)
```

<!-- ### AIC -->

<!-- ```{r} -->
<!-- # interaction model better by AIC -->
<!-- AIC(gam_NA_interaction, gam_NA_main) -->
<!-- ``` -->

### Plot smooths

```{r}
# how to make better plots?
plot(gam_na, scheme = 2
     ,page = 1
     ,shift = coef(gam_na)[1]
     )
```

### Plot predicted log(NO3)

```{r}
vis.gam(gam_na, view = c("ECO_TYPE", "log_DOC_med"), theta = 300)
vis.gam(gam_na, view = c("ECO_TYPE", "log_phos_med"), theta = 300)#330)

```

### Check k

Model assumptions not met. Even ignoring issues with spatial correlation. This may be problematic since GAM can be more sensitive to misspecification than GLM according to mgcv docs.

Residuals are right skewed (see qq plot and histogram). Also, appears to show heteroskedasticity, low predicted values have higher error than high predicted values (see residual vs fitted plot).

Values of basis dimension k are ok since edf < k.

```{r}
# need to fix this problem
# gam.check(gam_NA_interaction)
gam.check(gam_na)
```