---
title: "interactive CNP plot"
author: "David Nguyen"
date: '2022-04-08'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(plotly)
library(tidyverse)
setwd("C://Users//david//Documents//STOICH//STOICH.Aim1//Code/DNguyen")
```


# 3D plots

These interactive plots shows measurements of:

* DOC
* NO3 as N
* PO4 as P

NO3 is our response and DOC and PO4 are possible explanatory variables.

In the first plot (no transformations), it is clear that there is a "hockey-stick" type of relationship between NO3 ~ DOC*PO4. This indicates that modeling relative changes could be a useful approach, e.g., taking logs.

The second plot shows how log transformation linearizes this relationship.

```{r message = FALSE, warning = FALSE}
ALL_CNP <- read.csv("ALL_CNP.csv")

fig <- plot_ly(ALL_CNP, x = ~DOC, z = ~NO3.as.N, y = ~PO4.as.P, color = ~ECO_TYPE, colors = c('#BF382A', '#0C4B8E'))

fig <- fig %>% add_markers(opacity = 0.3)

fig

fig_log <- fig %>% layout(scene = list(xaxis = list(title = 'DOC', type = "log"),

                     zaxis = list(title = 'NO3', type = "log"),

                     yaxis = list(title = 'PO4', type = "log")))


fig_log
```


# NA

### Rivers

```{r}
ALL_CNP_med <- read.csv("data/ALL_CNP_med.csv")

cnp_na_rivers <- ALL_CNP_med %>% filter(ECO_TYPE == "River/Stream", region == "North America")

fig <- plot_ly(cnp_na_rivers, x = ~DOC_med, z = ~N_med, y = ~P_med, color = ~ECO_TYPE, colors = c('#BF382A', '#0C4B8E'))

fig <- fig %>% add_markers(opacity = 0.3)

fig

fig_log <- fig %>% layout(scene = list(xaxis = list(title = 'DOC', type = "log"),

                     zaxis = list(title = 'NO3', type = "log"),

                     yaxis = list(title = 'PO4', type = "log")))


fig_log
```

### Lakes

```{r}
cnp_na_lakes <- ALL_CNP_med %>% filter(ECO_TYPE == "Lake", region == "North America")

fig <- plot_ly(cnp_na_lakes, x = ~DOC_med, z = ~N_med, y = ~TP_med, color = ~ECO_TYPE, colors = c('#BF382A', '#0C4B8E'))

fig <- fig %>% add_markers(opacity = 0.3)

fig

fig_log <- fig %>% layout(scene = list(xaxis = list(title = 'DOC', type = "log"),

                     zaxis = list(title = 'NO3', type = "log"),

                     yaxis = list(title = 'PO4', type = "log")))


fig_log
```

# EU
### Rivers

``` {r }
cnp_eu_rivers <- ALL_CNP_med %>% filter(ECO_TYPE == "River/Stream", region == "Europe")

fig <- plot_ly(cnp_eu_rivers, x = ~DOC_med, z = ~N_med, y = ~P_med, color = ~ECO_TYPE, colors = c('#BF382A', '#0C4B8E'))

fig <- fig %>% add_markers(opacity = 0.3)

fig

fig_log <- fig %>% layout(scene = list(xaxis = list(title = 'DOC', type = "log"),

                     zaxis = list(title = 'NO3', type = "log"),

                     yaxis = list(title = 'PO4', type = "log")))


fig_log
```


### Lakes

```{r}
cnp_eu_lakes <- ALL_CNP_med %>% filter(ECO_TYPE == "Lake", region == "Europe")

fig <- plot_ly(cnp_eu_lakes, x = ~DOC_med, z = ~N_med, y = ~TP_med, color = ~ECO_TYPE, colors = c('#BF382A', '#0C4B8E'))

fig <- fig %>% add_markers(opacity = 0.3)

fig

fig_log <- fig %>% layout(scene = list(xaxis = list(title = 'DOC', type = "log"),

                     zaxis = list(title = 'NO3', type = "log"),

                     yaxis = list(title = 'PO4', type = "log")))


fig_log
```