---
title: "make prediction plots"
author: "David Nguyen"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(mgcv)
library(varycoef)

# Get average regression relationship averaged over space
# Similar to predict.lm
# get regression line and se conditioned on all RE == 0
predict_SVC <- function(svc.mod, Xp) {

  # Get vcov of fixed effects
  Sigma_final <- svc.mod$MLE$comp.args$Sigma_final
  X <- svc.mod$MLE$call.args$X
  Sigma_FE <- solve(crossprod(X, solve(Sigma_final, X)))
  
  # get predicted values
  fit <- Xp %*% coef(svc.mod)

  # SE = X %*% V %*% t(X)
  var.fit <- rowSums((Xp %*% Sigma_FE) * Xp)
  se.fit <- sqrt(var.fit)

  # return output as list
  return(list("fit" = fit, "se.fit" = se.fit))
}

# read in data
# original data
cnp <- read.csv("data/ALL_CNP_med.csv")

cnp$ECO_TYPE <- factor(cnp$ECO_TYPE) # for fitting gam by s(.., by = ECO_TYPE)

# merge median of medians of PO4 (lakes) and TP (lakes) into one variable
# this is so we can fit a common model for lakes and lakes
cnp <- cnp %>% mutate(phos_med = ifelse(ECO_TYPE == "River/Stream", 
                                                 P_med,
                                                 TP_med))
# appropriate P measurment missing for 5 EU lakes
# cnp_med %>% group_by(ECO_TYPE, region) %>% summarise(missing_phos = sum(is.na(phos_med)))
cnp <- cnp %>% filter(!is.na(phos_med))

# log everything
cnp <- cnp %>% mutate(log_N_med = log(N_med),
                              log_phos_med = log(phos_med),
                              log_DOC_med = log(DOC_med))


# split between EU and NA, lakes and lakes
cnp_eu_river <- cnp %>% filter(ECO_TYPE == "River/Stream", region == "Europe")
cnp_eu_lake <- cnp %>% filter(ECO_TYPE == "Lake", region == "Europe") %>%# 
  filter(!is.na(TP_med))

cnp_na_river <- cnp %>% filter(ECO_TYPE == "River/Stream", region == "North America")
cnp_na_lake <- cnp %>% filter(ECO_TYPE == "Lake", region == "North America")

# read in fitted models
# SVC
svc_na_lake <- readRDS("models/svc_na_lake_tp.RDS")
svc_na_river <- readRDS("models/svc_na_river.RDS")
svc_eu_lake <- readRDS("models/svc_eu_lake_tp.RDS")
svc_eu_river <- readRDS("models/svc_eu_river.RDS")


# GAM
gam_eu <- readRDS("models/gam_eu.RDS")
gam_na <- readRDS("models/gam_na.RDS")

# OLS
ols_eu <- lm(log_N_med ~ ECO_TYPE * log_DOC_med + ECO_TYPE * log_phos_med + ECO_TYPE + log_DOC_med*log_phos_med, 
             data = filter(cnp, region == "Europe") )
ols_na <- lm(log_N_med ~ ECO_TYPE * log_DOC_med + ECO_TYPE * log_phos_med + ECO_TYPE + log_DOC_med*log_phos_med, 
             data = filter(cnp, region == "North America") )
```

# Prediction grid

First need to choose what values to predict mean log median of medians nitrate level.
We want to make sure we aren't making prediction at DOC/P values that are very sparse.
We will use a grid of values between 3 - 97% percentiles of DOC while P is held at the median, and vice versa. Note, it would be better to use more values than just the median, but it is hard to show in a plot.

Using these percentiles were determined subjectivelly by choosing what seemed to be a good trade-off between predicting over a large interval of values while keeping to concentrations that are frequenty observed. See joint_dist_cp.pdf file in google drive for additional details.

The quantiles will be determined for each subset, e.g., lakes in north america, not quantiles for the entire data set.
This means that model predictions for different subsets of the data will not overlap the same values of DOC/P.


```{r make_new_data, eval = FALSE}
# chose quantile
fix_q <-  0.75 # adjust this value to get different quantiles to fix doc/p at. eg. 0.5 for median.

# DOC quantiles to use
q_doc_eu_lake <- quantile(cnp_eu_lake$log_DOC_med, fix_q)
q_doc_eu_river <- quantile(cnp_eu_river$log_DOC_med, fix_q)
q_doc_na_lake <- quantile(cnp_na_lake$log_DOC_med, fix_q)
q_doc_na_river <- quantile(cnp_na_river$log_DOC_med, fix_q)

# Phos quantiles to use
q_phos_eu_lake <- quantile(cnp_eu_lake$log_phos_med, fix_q)
q_phos_eu_river <- quantile(cnp_eu_river$log_phos_med, fix_q)
q_phos_na_lake <- quantile(cnp_na_lake$log_phos_med, fix_q)
q_phos_na_river <- quantile(cnp_na_river$log_phos_med, fix_q)

# DOC bounds
b_doc_eu_lake <- quantile(cnp_eu_lake$log_DOC_med, c(0.03, 0.97))
b_doc_eu_river <- quantile(cnp_eu_river$log_DOC_med, c(0.03, 0.97))
b_doc_na_lake <- quantile(cnp_na_lake$log_DOC_med, c(0.03, 0.97))
b_doc_na_river <- quantile(cnp_na_river$log_DOC_med, c(0.03, 0.97))

# phos bounds
b_phos_eu_lake <- quantile(cnp_eu_lake$log_phos_med, c(0.03, 0.97))
b_phos_eu_river <- quantile(cnp_eu_river$log_phos_med, c(0.03, 0.97))
b_phos_na_lake <- quantile(cnp_na_lake$log_phos_med, c(0.03, 0.97))
b_phos_na_river <- quantile(cnp_na_river$log_phos_med, c(0.03, 0.97))

# make df for varying DOC
nd_eu_lake_doc <- expand.grid(log_DOC_med = seq(b_doc_eu_lake[1], b_doc_eu_lake[2], length = 100),
                              log_phos_med = q_phos_eu_lake)

nd_eu_river_doc <- expand.grid(log_DOC_med = seq(b_doc_eu_river[1], b_doc_eu_river[2], length = 100),
                              log_phos_med = q_phos_eu_river)

nd_na_lake_doc <- expand.grid(log_DOC_med = seq(b_doc_na_lake[1], b_doc_na_lake[2], length = 100),
                              log_phos_med = q_phos_na_lake)

nd_na_river_doc <- expand.grid(log_DOC_med = seq(b_doc_na_river[1], b_doc_na_river[2], length = 100),
                              log_phos_med = q_phos_na_river)

# make df for varying DOC
nd_eu_lake_phos <- expand.grid(log_DOC_med = q_doc_eu_lake,
                              log_phos_med = seq(b_phos_eu_lake[1], b_phos_eu_lake[2], length = 100))

nd_eu_river_phos <- expand.grid(log_DOC_med = q_doc_eu_river,
                              log_phos_med = seq(b_phos_eu_river[1], b_phos_eu_river[2], length = 100))

nd_na_lake_phos <- expand.grid(log_DOC_med = q_doc_na_lake,
                              log_phos_med = seq(b_phos_na_lake[1], b_phos_na_lake[2], length = 100))

nd_na_river_phos <- expand.grid(log_DOC_med = q_doc_na_river,
                              log_phos_med = seq(b_phos_na_river[1], b_phos_na_river[2], length = 100))

# combine eco_types within region for OLS and GAM
nd_eu_doc <- rbind(nd_eu_lake_doc %>% mutate(ECO_TYPE = "Lake"),  
                   nd_eu_river_doc %>% mutate(ECO_TYPE = "River/Stream"))

nd_eu_phos <- rbind(nd_eu_lake_phos %>% mutate(ECO_TYPE = "Lake"),  
                   nd_eu_river_phos %>% mutate(ECO_TYPE = "River/Stream"))

nd_na_doc <- rbind(nd_na_lake_doc %>% mutate(ECO_TYPE = "Lake"),  
                   nd_na_river_doc %>% mutate(ECO_TYPE = "River/Stream"))

nd_na_phos <- rbind(nd_na_lake_phos %>% mutate(ECO_TYPE = "Lake"),  
                   nd_na_river_phos %>% mutate(ECO_TYPE = "River/Stream"))

# make model frame for SVC model
# 4 columns: intercept, logDOC, logPhos, interaction

# vary DOC
svc_data_eu_lake_doc <- cbind(1, nd_eu_lake_doc, nd_eu_lake_doc[,1] * nd_eu_lake_doc[,2])
svc_data_eu_river_doc <- cbind(1, nd_eu_river_doc, nd_eu_river_doc[,1] * nd_eu_river_doc[,2])

svc_data_na_lake_doc <- cbind(1, nd_na_lake_doc, nd_na_lake_doc[,1] * nd_na_lake_doc[,2])
svc_data_na_river_doc <- cbind(1, nd_na_river_doc, nd_na_river_doc[,1] * nd_na_river_doc[,2])

# vary PHOS
svc_data_eu_lake_phos <- cbind(1, nd_eu_lake_phos, nd_eu_lake_phos[,1] * nd_eu_lake_phos[,2])
svc_data_eu_river_phos <- cbind(1, nd_eu_river_phos, nd_eu_river_phos[,1] * nd_eu_river_phos[,2])

svc_data_na_lake_phos <- cbind(1, nd_na_lake_phos, nd_na_lake_phos[,1] * nd_na_lake_phos[,2])
svc_data_na_river_phos <- cbind(1, nd_na_river_phos, nd_na_river_phos[,1] * nd_na_river_phos[,2])
```

```{r predictions_OLS, eval = FALSE}
# predictions for each region
# doc varies
pred_eu_doc <- 
  cbind(nd_eu_doc, predict(ols_eu, newdata = nd_eu_doc, interval = "confidence")) %>%
  # mutate(region = "Europe") %>% 
  rename(fit_ols = fit, lwr_ols = lwr, upr_ols = upr)
pred_na_doc <- 
  cbind(nd_na_doc, predict(ols_na, newdata = nd_na_doc, interval = "confidence"))%>%
  # mutate(region = "North America") %>% 
  rename(fit_ols = fit, lwr_ols = lwr, upr_ols = upr)

# phos varies
pred_eu_phos <- 
  cbind(nd_eu_phos, predict(ols_eu, newdata = nd_eu_phos, interval = "confidence")) %>%
  # mutate(region = "Europe") %>% 
  rename(fit_ols = fit, lwr_ols = lwr, upr_ols = upr)
pred_na_phos <- 
  cbind(nd_na_phos, predict(ols_na, newdata = nd_na_phos, interval = "confidence")) %>%
  # mutate(region = "North America") %>% 
  rename(fit_ols = fit, lwr_ols = lwr, upr_ols = upr)

```

```{r eval = FALSE}
# pred EU
pred_gam_eu_doc <- predict(gam_eu, newdata = nd_eu_doc, type = "response", se.fit = TRUE)
pred_gam_eu_phos <- predict(gam_eu, newdata = nd_eu_phos, type = "response", se.fit = TRUE)

# pred NA
pred_gam_na_doc <- predict(gam_na, newdata = nd_na_doc, type = "response", se.fit = TRUE)
pred_gam_na_phos <- predict(gam_na, newdata = nd_na_phos, type = "response", se.fit = TRUE)

# add prediction and CI to data frame
pred_eu_doc$fit_gam <- pred_gam_eu_doc$fit
pred_eu_doc$lwr_gam <- pred_gam_eu_doc$fit - 2 * pred_gam_eu_doc$se.fit
pred_eu_doc$upr_gam <- pred_gam_eu_doc$fit + 2 * pred_gam_eu_doc$se.fit

pred_eu_phos$fit_gam <- pred_gam_eu_phos$fit
pred_eu_phos$lwr_gam <- pred_gam_eu_phos$fit - 2 * pred_gam_eu_phos$se.fit
pred_eu_phos$upr_gam <- pred_gam_eu_phos$fit + 2 * pred_gam_eu_phos$se.fit

pred_na_doc$fit_gam <- pred_gam_na_doc$fit
pred_na_doc$lwr_gam <- pred_gam_na_doc$fit - 2 * pred_gam_na_doc$se.fit
pred_na_doc$upr_gam <- pred_gam_na_doc$fit + 2 * pred_gam_na_doc$se.fit

pred_na_phos$fit_gam <- pred_gam_na_phos$fit
pred_na_phos$lwr_gam <- pred_gam_na_phos$fit - 2 * pred_gam_na_phos$se.fit
pred_na_phos$upr_gam <- pred_gam_na_phos$fit + 2 * pred_gam_na_phos$se.fit
```

```{r eval = FALSE}
pred_svc_eu_doc_lake <- predict_SVC(svc_eu_lake, Xp = as.matrix(svc_data_eu_lake_doc))
pred_svc_na_doc_lake <- predict_SVC(svc_na_lake, Xp = as.matrix(svc_data_na_lake_doc))

pred_svc_eu_doc_river <- predict_SVC(svc_eu_river, Xp = as.matrix(svc_data_eu_river_doc))
pred_svc_na_doc_river <- predict_SVC(svc_na_river, Xp = as.matrix(svc_data_na_river_doc))

pred_svc_eu_phos_lake <- predict_SVC(svc_eu_lake, Xp = as.matrix(svc_data_eu_lake_phos))
pred_svc_na_phos_lake <- predict_SVC(svc_na_lake, Xp = as.matrix(svc_data_na_lake_phos))

pred_svc_eu_phos_river <- predict_SVC(svc_eu_river, Xp = as.matrix(svc_data_eu_river_phos))
pred_svc_na_phos_river <- predict_SVC(svc_na_river, Xp = as.matrix(svc_data_na_river_phos))

# add prediction and CI to data frame
# EU doc
pred_eu_doc$fit_svc <- c(pred_svc_eu_doc_lake$fit, pred_svc_eu_doc_river$fit) 
pred_eu_doc$lwr_svc <- pred_eu_doc$fit_svc - 2 * c(pred_svc_eu_doc_lake$fit, pred_svc_eu_doc_river$fit) 
pred_eu_doc$upr_svc <- pred_eu_doc$fit_svc + 2 * c(pred_svc_eu_doc_lake$fit, pred_svc_eu_doc_river$fit)

# NA doc
pred_na_doc$fit_svc <- c(pred_svc_na_doc_lake$fit, pred_svc_na_doc_river$fit) 
pred_na_doc$lwr_svc <- pred_na_doc$fit_svc - 2 * c(pred_svc_na_doc_lake$fit, pred_svc_na_doc_river$fit) 
pred_na_doc$upr_svc <- pred_na_doc$fit_svc + 2 * c(pred_svc_na_doc_lake$fit, pred_svc_na_doc_river$fit)

# EU doc
pred_eu_phos$fit_svc <- c(pred_svc_eu_phos_lake$fit, pred_svc_eu_phos_river$fit) 
pred_eu_phos$lwr_svc <- pred_eu_phos$fit_svc - 2 * c(pred_svc_eu_phos_lake$fit, pred_svc_eu_phos_river$fit) 
pred_eu_phos$upr_svc <- pred_eu_phos$fit_svc + 2 * c(pred_svc_eu_phos_lake$fit, pred_svc_eu_phos_river$fit)

# NA doc
pred_na_phos$fit_svc <- c(pred_svc_na_phos_lake$fit, pred_svc_na_phos_river$fit) 
pred_na_phos$lwr_svc <- pred_na_phos$fit_svc - 2 * c(pred_svc_na_phos_lake$fit, pred_svc_na_phos_river$fit) 
pred_na_phos$upr_svc <- pred_na_phos$fit_svc + 2 * c(pred_svc_na_phos_lake$fit, pred_svc_na_phos_river$fit)

```

```{r eval = FALSE}
# combine regions
pred_doc <- rbind(pred_na_doc %>% mutate(region = "North America"), pred_eu_doc %>% mutate(region = "Europe"))
pred_phos <- rbind(pred_na_phos %>% mutate(region = "North America"), pred_eu_phos %>% mutate(region = "Europe"))

pred_doc$fix_q <- fix_q
pred_phos$fix_q <- fix_q

pred_doc_name <- paste0("pred_doc_", fix_q * 100, ".RDS")
pred_phos_name <- paste0("pred_phos_", fix_q * 100, ".RDS")

saveRDS(pred_doc, file = pred_doc_name)
saveRDS(pred_phos, file = pred_phos_name)
```


# plots

```{r}
# load prediction which were computed earlier
# vary doc
pred_doc_25 <- readRDS("pred_doc_25.RDS")
pred_doc_50 <- readRDS("pred_doc_50.RDS")
pred_doc_75 <- readRDS("pred_doc_75.RDS")
# vary phos
pred_phos_25 <- readRDS("pred_phos_25.RDS")
pred_phos_50 <- readRDS("pred_phos_50.RDS")
pred_phos_75 <- readRDS("pred_phos_75.RDS")

pred_doc <- rbind(pred_doc_25, pred_doc_50, pred_doc_75)
pred_phos <- rbind(pred_phos_25, pred_phos_50, pred_phos_75)
```


Color legend: SVC = darkgreen, OLS = steelblue, GAM = goldenrod

When DOC is varied, P is held at median and vice-versa.

Interpretion: remember, we did log-log regression, so the slope of the curve is what corresponds to the exponential decay parameter k in Taylor and Townsend. So, if the slope of the curve is negative, there is a exponential decaying relationship and if the slope of the curve is positive, there is an exponential increasing relationship.

Notice that for the GAM, we can see that some values of DOC/P concentration are positive whereas others are negative, Scientifically, this suggests that the concentration of a chemical can affect whether it is associated with higher or lower nitrate.

### All quantiles

```{r}
pred_doc %>%
  ggplot(aes(x = log_DOC_med, y = fit_ols, group = factor(fix_q), linetype = factor(fix_q))) +
  geom_line(aes(y = fit_svc), col = "darkgreen") +
  geom_line(aes(y = fit_ols), col = "steelblue") + 
  geom_line(aes(y = fit_gam), col = "goldenrod") +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.05, fill = "darkgreen") +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.05, fill = "steelblue") +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.05, fill = "goldenrod") +
  facet_grid(region~ECO_TYPE) +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(DOC mg/l))",
       linetype = "quantile P")
```

```{r}
pred_phos %>%
  ggplot(aes(x = log_phos_med, group = factor(fix_q), linetype = factor(fix_q))) +
  geom_line(aes(y = fit_svc), col = "darkgreen") +
  geom_line(aes(y = fit_ols), col = "steelblue") + 
  geom_line(aes(y = fit_gam), col = "goldenrod") +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.05, fill = "darkgreen") +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.05, fill = "steelblue") +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.05, fill = "goldenrod") +
  facet_grid(region~ECO_TYPE) +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(P mg/l))",
       linetype = "quantile DOC")
```

# 25th quantile

### All axis on same scale

```{r}
pred_doc_25 %>%
  pivot_longer(cols = starts_with("fit"),
               names_sep = "_",
               names_to = c("fit", "method"),
               values_to = "value") %>%
  ggplot(aes(x = log_DOC_med, fill = method)) +
  geom_line(aes(y = value,linetype = method, col = method)) +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.1) +
  facet_grid(region~ECO_TYPE) +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(DOC mg/l))")

pred_phos_25 %>%
  pivot_longer(cols = starts_with("fit"),
               names_sep = "_",
               names_to = c("fit", "method"),
               values_to = "value") %>%
  ggplot(aes(x = log_phos_med, fill = method)) +
  geom_line(aes(y = value,linetype = method, col = method)) +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.1) +
  facet_grid(region~ECO_TYPE) +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(P mg/l))")

```

### Axis on different scales

```{r}
pred_doc_25 %>%
  pivot_longer(cols = starts_with("fit"),
               names_sep = "_",
               names_to = c("fit", "method"),
               values_to = "value") %>%
  ggplot(aes(x = log_DOC_med, fill = method)) +
  geom_line(aes(y = value,linetype = method, col = method)) +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.1) +
  facet_grid(region~ECO_TYPE, scales = "free") +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(DOC mg/l))")

pred_phos_25 %>%
  pivot_longer(cols = starts_with("fit"),
               names_sep = "_",
               names_to = c("fit", "method"),
               values_to = "value") %>%
  ggplot(aes(x = log_phos_med, fill = method)) +
  geom_line(aes(y = value,linetype = method, col = method)) +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.1) +
  facet_grid(region~ECO_TYPE, scales = "free") +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(P mg/l))")

```


```{r eval = FALSE}
pred_doc_25 %>%
  ggplot(aes(x = log_DOC_med)) +
  geom_line(aes(y = fit_svc), col = "darkgreen") +
  geom_line(aes(y = fit_ols), col = "steelblue") + 
  geom_line(aes(y = fit_gam), col = "goldenrod") +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.1, fill = "darkgreen") +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.1, fill = "steelblue") +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.1, fill = "goldenrod") +
  facet_grid(region~ECO_TYPE) +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(DOC mg/l))")

pred_phos_25 %>%
  ggplot(aes(x = log_phos_med)) +
  geom_line(aes(y = fit_svc), col = "darkgreen") +
  geom_line(aes(y = fit_ols), col = "steelblue") + 
  geom_line(aes(y = fit_gam), col = "goldenrod") +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.1, fill = "darkgreen") +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.1, fill = "steelblue") +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.1, fill = "goldenrod") +
  facet_grid(region~ECO_TYPE) +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(P mg/l))")

```

# 50th quantile

### All axis on same scale

```{r}
pred_doc_50 %>%
  pivot_longer(cols = starts_with("fit"),
               names_sep = "_",
               names_to = c("fit", "method"),
               values_to = "value") %>%
  ggplot(aes(x = log_DOC_med, fill = method)) +
  geom_line(aes(y = value,linetype = method, col = method)) +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.1) +
  facet_grid(region~ECO_TYPE) +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(DOC mg/l))")

pred_phos_50 %>%
  pivot_longer(cols = starts_with("fit"),
               names_sep = "_",
               names_to = c("fit", "method"),
               values_to = "value") %>%
  ggplot(aes(x = log_phos_med, fill = method)) +
  geom_line(aes(y = value,linetype = method, col = method)) +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.1) +
  facet_grid(region~ECO_TYPE) +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(P mg/l))")

```

### Axis on different scales

```{r}
pred_doc_50 %>%
  pivot_longer(cols = starts_with("fit"),
               names_sep = "_",
               names_to = c("fit", "method"),
               values_to = "value") %>%
  ggplot(aes(x = log_DOC_med, fill = method)) +
  geom_line(aes(y = value,linetype = method, col = method)) +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.1) +
  facet_grid(region~ECO_TYPE, scales = "free") +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(DOC mg/l))")

pred_phos_50 %>%
  pivot_longer(cols = starts_with("fit"),
               names_sep = "_",
               names_to = c("fit", "method"),
               values_to = "value") %>%
  ggplot(aes(x = log_phos_med, fill = method)) +
  geom_line(aes(y = value,linetype = method, col = method)) +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.1) +
  facet_grid(region~ECO_TYPE, scales = "free") +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(P mg/l))")

```

```{r eval = FALSE}
pred_doc_50 %>%
  ggplot(aes(x = log_DOC_med)) +
  geom_line(aes(y = fit_svc), col = "darkgreen") +
  geom_line(aes(y = fit_ols), col = "steelblue") + 
  geom_line(aes(y = fit_gam), col = "goldenrod") +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.1, fill = "darkgreen") +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.1, fill = "steelblue") +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.1, fill = "goldenrod") +
  facet_grid(region~ECO_TYPE) +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(DOC mg/l))")

pred_phos_50 %>%
  ggplot(aes(x = log_phos_med)) +
  geom_line(aes(y = fit_svc), col = "darkgreen") +
  geom_line(aes(y = fit_ols), col = "steelblue") + 
  geom_line(aes(y = fit_gam), col = "goldenrod") +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.1, fill = "darkgreen") +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.1, fill = "steelblue") +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.1, fill = "goldenrod") +
  facet_grid(region~ECO_TYPE) +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(P mg/l))")

```

# 75th quantile

### All axis on same scale

```{r eval = FALSE}
pred_doc_75 %>%
  ggplot(aes(x = log_DOC_med)) +
  geom_line(aes(y = fit_svc), col = "darkgreen") +
  geom_line(aes(y = fit_ols), col = "steelblue") + 
  geom_line(aes(y = fit_gam), col = "goldenrod") +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.1, fill = "darkgreen") +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.1, fill = "steelblue") +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.1, fill = "goldenrod") +
  facet_grid(region~ECO_TYPE) +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(DOC mg/l))")

pred_phos_75 %>%
  ggplot(aes(x = log_phos_med)) +
  geom_line(aes(y = fit_svc), col = "darkgreen") +
  geom_line(aes(y = fit_ols), col = "steelblue") + 
  geom_line(aes(y = fit_gam), col = "goldenrod") +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.1, fill = "darkgreen") +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.1, fill = "steelblue") +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.1, fill = "goldenrod") +
  facet_grid(region~ECO_TYPE) +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(P mg/l))")

```

### Axis on different scales

```{r}
pred_doc_75 %>%
  pivot_longer(cols = starts_with("fit"),
               names_sep = "_",
               names_to = c("fit", "method"),
               values_to = "value") %>%
  ggplot(aes(x = log_DOC_med, fill = method)) +
  geom_line(aes(y = value,linetype = method, col = method)) +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.1) +
  facet_grid(region~ECO_TYPE, scales = "free") +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(DOC mg/l))")

pred_phos_75 %>%
  pivot_longer(cols = starts_with("fit"),
               names_sep = "_",
               names_to = c("fit", "method"),
               values_to = "value") %>%
  ggplot(aes(x = log_phos_med, fill = method)) +
  geom_line(aes(y = value,linetype = method, col = method)) +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.1) +
  facet_grid(region~ECO_TYPE, scales = "free") +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(P mg/l))")

```

```{r}
pred_doc_75 %>%
  pivot_longer(cols = starts_with("fit"),
               names_sep = "_",
               names_to = c("fit", "method"),
               values_to = "value") %>%
  ggplot(aes(x = log_DOC_med, fill = method)) +
  geom_line(aes(y = value,linetype = method, col = method)) +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.1) +
  facet_grid(region~ECO_TYPE) +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(DOC mg/l))")

pred_phos_75 %>%
  pivot_longer(cols = starts_with("fit"),
               names_sep = "_",
               names_to = c("fit", "method"),
               values_to = "value") %>%
  ggplot(aes(x = log_phos_med, fill = method)) +
  geom_line(aes(y = value,linetype = method, col = method)) +
  geom_ribbon(aes(ymin = lwr_svc, ymax = upr_svc) ,alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_ols, ymax = upr_ols), alpha = 0.1) +
  geom_ribbon(aes(ymin = lwr_gam, ymax = upr_gam), alpha = 0.1) +
  facet_grid(region~ECO_TYPE) +
  labs(y = "Predicted log(med(N mg/l)) 95% CI",
       x = "log(med(P mg/l))")

```


# Caveats

Not sure what's going on with the SVC confidence intervals. They seem weirdly narrow at some values.

The SVC spatial average prediction doesn't seem to match well with the GAM and OLS. I think a possible reason for this, is that when I get this prediction, I set all random effects equal to zero instead of marginalizing over them.

The GAM model I'm plotting does not include a spatial smooth. This is because there is no built-in predict method for predicting the response when the spatial smooth is set to zero. So it will have the same issues wrt to underestimation of uncertainty as the OLS model.

