---
title: "SVC maps"
author: "David Nguyen"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(varycoef)
library(viridis)

# make maps of sd(SVC)
# svc_pred_sd: data frame of sd(svc)
# which_svc: choose beta coefficient (0L = intercept, 1L = DOC slope, 2L = P slope)
make_svc_sd_map <- function(svc_pred_sd, which_svc = NA){
  
  if(!(which_svc %in% c(0L,1L,2L))) stop("which_svc must be either 0L, 1L, 2L")
  
  state_map <- ggplot2::map_data("state")
  world_map <- ggplot2::map_data("world")
  
  if (which_svc == 0L) {
   p <-  svc_pred_sd %>%
  ggplot() + 
    geom_tile(aes(
      x = LON ,
      y = LAT ,
      fill = SVC_1_sd
    ), col = NA, alpha = 1) +
    geom_map(aes(map_id = region), col = "gray", fill = NA, 
           data = state_map, map = state_map) +
  geom_map(aes(map_id = region), col = "gray", fill = NA, 
           data = world_map, map = world_map) +
    labs(#title = "Uncertainty of spatially varying intercept", 
         x = "Longitude", y = "Latitude"
         ,fill = expression(sd(widehat(beta[0] + beta[0](s))))
         ) +
  coord_fixed() +
    scale_fill_viridis() +
  theme(legend.position = "top")
    
  } else  if ( which_svc == 1L) {
   p <-   svc_pred_sd %>%
  ggplot() + 
    geom_tile(aes(
      x = LON ,
      y = LAT ,
      fill = SVC_2_sd
    ), col = NA, alpha = 1) +
    geom_map(aes(map_id = region), col = "gray", fill = NA, 
           data = state_map, map = state_map) +
  geom_map(aes(map_id = region), col = "gray", fill = NA, 
           data = world_map, map = world_map) +
    labs(#title = "Uncertainty of spatially varying intercept", 
         x = "Longitude", y = "Latitude"
         ,fill = expression(sd(widehat(beta[1] + beta[1](s))))
         ) +
  coord_fixed() +
    scale_fill_viridis() +
  theme(legend.position = "top")
   
  }  else if( which_svc == 2L ) {
  p <-  svc_pred_sd %>%
  ggplot() + 
    geom_tile(aes(
      x = LON ,
      y = LAT ,
      fill = SVC_3_sd
    ), col = NA, alpha = 1) +
    geom_map(aes(map_id = region), col = "gray", fill = NA, 
           data = state_map, map = state_map) +
  geom_map(aes(map_id = region), col = "gray", fill = NA, 
           data = world_map, map = world_map) +
    labs(#title = "Uncertainty of spatially varying intercept", 
         x = "Longitude", y = "Latitude"
         ,fill = expression(sd(widehat(beta[2] + beta[2](s))))
         ) +
  coord_fixed() +
    scale_fill_viridis() +
  theme(legend.position = "top")
  }
  print(p)
}


# make maps of SVC
# svc_pred: data frame of sd(svc)
# which_svc: choose beta coefficient (0L = intercept, 1L = DOC slope, 2L = P slope)
make_svc_map <- function(svc_pred, which_svc = NA){
  
  if(!(which_svc %in% c(0L,1L,2L))) stop("which_svc must be either 0L, 1L, 2L")
  
  state_map <- ggplot2::map_data("state")
  world_map <- ggplot2::map_data("world")
  
  if (which_svc == 0L) {
   p <-  svc_pred %>%
  ggplot() + 
    geom_tile(aes(
      x = loc_1 ,
      y = loc_2 ,
      fill = SVC_1
    ), col = NA, alpha = 1) +
    geom_map(aes(map_id = region), col = "gray", fill = NA, 
           data = state_map, map = state_map) +
  geom_map(aes(map_id = region), col = "gray", fill = NA, 
           data = world_map, map = world_map) +
    labs(#title = "Uncertainty of spatially varying intercept", 
         x = "Longitude", y = "Latitude"
         ,fill = expression(widehat(beta[0] + beta[0](s)))
         ) +
  coord_fixed() +
    scale_fill_viridis() +
  theme(legend.position = "top")
    
  } else  if ( which_svc == 1L) {
   p <-   svc_pred %>%
  ggplot() + 
    geom_tile(aes(
      x = loc_1 ,
      y = loc_2 ,
      fill = SVC_2
    ), col = NA, alpha = 1) +
    geom_map(aes(map_id = region), col = "gray", fill = NA, 
           data = state_map, map = state_map) +
  geom_map(aes(map_id = region), col = "gray", fill = NA, 
           data = world_map, map = world_map) +
    labs(#title = "Uncertainty of spatially varying intercept", 
         x = "Longitude", y = "Latitude"
         ,fill = expression(widehat(beta[1] + beta[1](s)))
         ) +
  coord_fixed() +
    scale_fill_viridis() +
  theme(legend.position = "top")
   
  } else  if ( which_svc == 2L){
  p <-  svc_pred %>%
  ggplot() + 
    geom_tile(aes(
      x = loc_1 ,
      y = loc_2 ,
      fill = SVC_3
    ), col = NA, alpha = 1) +
    geom_map(aes(map_id = region), col = "gray", fill = NA, 
           data = state_map, map = state_map) +
  geom_map(aes(map_id = region), col = "gray", fill = NA, 
           data = world_map, map = world_map) +
    labs(#title = "Uncertainty of spatially varying intercept", 
         x = "Longitude", y = "Latitude"
         ,fill = expression(widehat(beta[2] + beta[2](s)))
         ) +
  coord_fixed() +
    scale_fill_viridis() +
  theme(legend.position = "top")
  }
  print(p)
}
  
# Calculate CI of individual SVC terms
# currenlty underestimates width of CI
# need to incorporate SE of fixed effect estimate

# returns a data frame combinining prediction, sd, lower and upper approximate 95% CI, 
# and whether slopes for DOC or P are significant and what direction effect is.
svc_ci <- function(pred_df, sd_df) {
  
  # join point estimate and sd of SVC terms
  # caluclate 95% CI for each SVC
  # might be wrong, doesn't account for se of fixed effect
  new_df <- 
  full_join(pred_df, 
          sd_df %>% rename(loc_1 = LON, loc_2 = LAT), 
          by = c("loc_1", "loc_2")) %>%
  mutate(lower_svc_1 = SVC_1 - 2 * SVC_1_sd,
         lower_svc_2 = SVC_2 - 2 * SVC_2_sd,
         lower_svc_3 = SVC_3 - 2 * SVC_3_sd,
         upper_svc_1 = SVC_1 + 2 * SVC_1_sd,
         upper_svc_2 = SVC_2 + 2 * SVC_2_sd,
         upper_svc_3 = SVC_3 + 2 * SVC_3_sd)

  # make variable for whether DOC or P slope is significantly diff from 0
  # based on whether previously calculated CI include 0 or not
  new_df <-
  new_df %>%
  mutate(doc_sig = case_when(lower_svc_2 > 0 ~ "+",
                              upper_svc_2 < 0 ~ "-",
                              TRUE ~ "n.s."),
         p_sig = case_when(lower_svc_3 > 0 ~ "+",
                              upper_svc_3 < 0 ~ "-",
                              TRUE ~ "n.s."))
return(new_df)
}

```

# Overview

Here, I produce maps of the individual spatially varying coefficients (SVC) for each of the 4 models (region X lake vs river) and 3 SVC terms (intercept (beta0), DOC slope (beta1), P slope (beta2))

I made 3 types of plots:

* plot of the point estimate of the SVC
* plots of the standard deviation (i.e., standard error) of the estimated svc
* plots of whether the estimate is significant and, if it is, what direction the slope is (positive or negative)


Note, I think the standard error estimates will be biased low since I have not included the uncertainty in the estimate of the fixed effect (i.e., the average slope averaged over space). This means CI will be too narrow and we may have elevated type 1 error. Also, non of the significance tests are corrected for multiple testing.

### NA LAKE

```{r}
# load SVC predictions and standard deviation of prediction
# these are calculated in calc_svc_sd.R
pred_na_lake <- readRDS("pred_na_lake.RDS")
pred_sd_na_lake <- readRDS("pred_sd_na_lake.RDS")
```

```{r svc_map_na_lake}
# make maps of predicted SVC
make_svc_map(pred_na_lake, 0L)  
make_svc_map(pred_na_lake, 1)  
make_svc_map(pred_na_lake, 2L)  
```

```{r sd_svc_map_na_lake}
# makes maps of standard deviation of predicted SVC
make_svc_sd_map(pred_sd_na_lake, 0L)  
make_svc_sd_map(pred_sd_na_lake, 1L)  
make_svc_sd_map(pred_sd_na_lake, 2L)  
```

```{r}
# calculate approx 95% CI
svc_na_lake_ci <- svc_ci(pred_na_lake, pred_sd_na_lake)

# plot maps of significance
svc_na_lake_ci %>%
  ggplot() +
  geom_tile(aes(x = loc_1, y = loc_2, fill = doc_sig)) +
  labs(title = "significance of DOC slope")

svc_na_lake_ci %>%
  ggplot() +
  geom_tile(aes(x = loc_1, y = loc_2, fill = p_sig)) +
  labs(title = "significance of P slope")
```

### NA RIVER

```{r }
# load SVC predictions and standard deviation of prediction
# these are calculated in calc_svc_sd.R
pred_na_river <- readRDS("pred_na_river.RDS")
pred_sd_na_river <- readRDS("pred_sd_na_river.RDS")
```

```{r svc_map_na_river}
# make maps of predicted SVC
make_svc_map(pred_na_river, 0L)  
make_svc_map(pred_na_river, 1L)  
make_svc_map(pred_na_river, 2L)  
```

```{r sd_svc_map_na_river}
# makes maps of standard deviation of predicted SVC
make_svc_sd_map(pred_sd_na_river, 0L)  
make_svc_sd_map(pred_sd_na_river, 1L)  
make_svc_sd_map(pred_sd_na_river, 2L)  
```

```{r}
# calculate approx 95% CI
svc_na_river_ci <- svc_ci(pred_na_river, pred_sd_na_river)

# plot maps of significance
svc_na_river_ci %>%
  ggplot() +
  geom_tile(aes(x = loc_1, y = loc_2, fill = doc_sig)) +
  labs(title = "significance of DOC slope")

svc_na_river_ci %>%
  ggplot() +
  geom_tile(aes(x = loc_1, y = loc_2, fill = p_sig)) +
  labs(title = "significance of P slope")
```

### EU LAKE

```{r}
# load SVC predictions and standard deviation of prediction
# these are calculated in calc_svc_sd.R
pred_eu_lake <- readRDS("pred_eu_lake.RDS")
pred_sd_eu_lake <- readRDS("pred_sd_eu_lake.RDS")
```

```{r svc_map_eu_lake}
# make maps of predicted SVC
make_svc_map(pred_eu_lake, 0L)  
make_svc_map(pred_eu_lake, 1L)  
make_svc_map(pred_eu_lake, 2L)  
```

```{r sd_svc_map_eu_lake}
# makes maps of standard deviation of predicted SVC
make_svc_sd_map(pred_sd_eu_lake, 0L)  
make_svc_sd_map(pred_sd_eu_lake, 1L)  
make_svc_sd_map(pred_sd_eu_lake, 2L)  
```

```{r}
# calculate approx 95% CI
svc_eu_lake_ci <- svc_ci(pred_eu_lake, pred_sd_eu_lake)

# plot maps of significance
svc_eu_lake_ci %>%
  ggplot() +
  geom_tile(aes(x = loc_1, y = loc_2, fill = doc_sig)) +
  labs(title = "significance of DOC slope")

svc_eu_lake_ci %>%
  ggplot() +
  geom_tile(aes(x = loc_1, y = loc_2, fill = p_sig)) +
  labs(title = "significance of P slope")
```

### EU RIVER

```{r }
# load SVC predictions and standard deviation of prediction
# these are calculated in calc_svc_sd.R
pred_eu_river <- readRDS("pred_eu_river.RDS")
pred_sd_eu_river <- readRDS("pred_sd_eu_river.RDS")
```

```{r svc_map_eu_river}
# make maps of predicted SVC
make_svc_map(pred_eu_river, 0L)  
make_svc_map(pred_eu_river, 1L)  
make_svc_map(pred_eu_river, 2L)  
```

```{r sd_svc_map_eu_river}
# makes maps of standard deviation of predicted SVC
make_svc_sd_map(pred_sd_eu_river, 0L)  
make_svc_sd_map(pred_sd_eu_river, 1L)  
make_svc_sd_map(pred_sd_eu_river, 2L)  
```

```{r}
# calculate approx 95% CI
svc_eu_river_ci <- svc_ci(pred_eu_river, pred_sd_eu_river)

# plot maps of significance
svc_eu_river_ci %>%
  ggplot() +
  geom_tile(aes(x = loc_1, y = loc_2, fill = doc_sig)) +
  labs(title = "significance of DOC slope")

svc_eu_river_ci %>%
  ggplot() +
  geom_tile(aes(x = loc_1, y = loc_2, fill = p_sig)) +
  labs(title = "significance of P slope")
```

