---
title: "OLS by country"
author: "David Nguyen"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(maps)
```

# Overview

```{r message=FALSE}
# load CNP data
cnp <- read_csv("https://raw.githubusercontent.com/swulfing/STOICH.Aim1/main/Data/ALL_CNP_April2022.csv")

# correcting DOC for LAGOS sites seems to make sense
# probably was a unit error
cnp <- cnp %>% mutate(DOC = ifelse(log10(DOC) < -1 & str_detect(SITE_ID, "LAGOS"),
                                   DOC * 10^3,
                                   DOC))

# get country for each site 
cnp <- 
  cnp %>% 
  mutate(country = map.where(database = "world", x = LON, y = LAT))

# number of countries
# cnp %>% pull(country) %>% unique() %>% length()

# take countries with at least 10 sites
country_subset <-
  cnp %>% 
  group_by(SITE_ID) %>% slice(1) %>% 
  group_by(country) %>% summarise(obs_country = n()) %>% filter(obs_country >= 10) %>%
  arrange(obs_country)

country_names <- country_subset$country
country_names <- country_names[!is.na(country_names)]

world_map <- map_data("world")

# get median of each SITE_ID over time
cnp_med <- 
  cnp %>% filter(country %in% country_names) %>%
  group_by(SITE_ID) %>%
  summarise(country = unique(country), 
            LAT = unique(LAT),
            LON = unique(LON),
            ECO_TYPE = unique(ECO_TYPE),
            N_med = median(NO3.as.N, na.rm = TRUE),
            P_med = median(PO4.as.P, na.rm = TRUE),
            DOC_med = median(DOC, na.rm = TRUE))

nobs_per_region <- 
  cnp_med %>% group_by(country) %>% 
  summarise(region = unique(country), n = n()) %>% 
  ungroup() %>% select(-country)
nobs_per_region$region[str_detect(nobs_per_region$region, "UK")] <- "UK"

mod_list <- vector("list", length = length(country_names))

for ( country_index in seq_along(country_names) ) {
  cnp_country <- cnp_med %>% filter(country == country_names[country_index])
  mod_list[[country_index]] <- lm(log10(N_med) ~ 1 + log10(DOC_med) + log10(P_med) 
                                  # + ECO_TYPE *log10(DOC_med) + ECO_TYPE *log10(P_med)
     , data = cnp_country)
}  

beta_country <- lapply(1:length(mod_list), function(x) coef(mod_list[[x]])) %>% bind_rows()
beta_country$region <- country_names
beta_country$region[str_detect(beta_country$region, "UK")] <- "UK"

beta_country_fill <- beta_country %>% pivot_longer(cols = names(coef(mod_list[[1]])),
                              names_to = "parameter",
                              values_to = "estimate")

# ggplot(beta_country_fill, aes(fill = estimate)) +
#   geom_map(map = world_subset,
#            aes(long, lat, map_id = region
#                )) +
#   coord_equal(ratio = 1) +
#   expand_limits() +
#     facet_wrap(~parameter)

# facet plot  
# beta_country_fill %>% filter(!str_detect(region, "US") ) %>% 
# ggplot() +
#   geom_map(aes(map_id = region, fill = estimate), map = world_map, col = "white") +
#   geom_point(data = cnp_med, aes(x = LON, y = LAT), alpha = 0.1, shape = 21) +
#   expand_limits(x = world_map$long, y = world_map$lat) +
#   viridis::scale_fill_viridis() +
#   coord_equal(xlim = c(-10, 30), ylim = c(35, 62.5)) +
#   facet_wrap(~parameter)
```

For a quick look at how the effects of $PO_4$ and $DOC$ affect $NO_3$ concentrations, we subsetted the CNP data set by country. This yielded `r length(unique(cnp$country))` total countries that had data. Of these countries we analyzed only countries with at least 10 sites which resulted in `r `length(country_names)` total countries in the following analysis.

For each country we fit the linear regression model using ordinary least squares: $log(NO_3)_i = \beta_0 + \beta_1 log(DOC)_i + \beta_2 log(PO_4)_i + \epsilon_i$

Where $i$ indexes a unique SITE_ID within each country and the response and covariates are log 10 transformed medians. ECO_TYPE (factor with levels: 1. River/stream; 2. Lake) was not included as covariate because both levels were not present in all subsets.

# Table of OLS coefficients

```{r}
left_join(beta_country, nobs_per_region, by = "region") %>% knitr::kable(caption = "OLS regression coefficients by country", digits = 3)
```


# Maps of regression OLS coefficients

The following maps show regression coefficients for European countries. The US was omitted from the map since there are is only 1 estimate for the US (see table 1.) Site locations are shown on the map by black circles.

There may be some spatial pattern in coefficients, but it is hard to tell since some regions have different numbers and spatial distribution of sites within each country. As such, by estimating country-specific models, some of the regression estimates are based on few observations (unstable estimates) or average over sites that may be very far apart. Also, it is possible that some of the variability in regression coefficients is due to omission of site ECO_TYPE which was found to have an interaction with the linear effect of $log(PO_4)$ and $log(DOC)$ in previous OLS regression over the entire data set.

```{r}
# individaul plots
# fill scale might be better since it is specific to each variable
beta_country %>%
  ggplot() +
  geom_map(aes(map_id = region, fill = `(Intercept)`), map = world_map, col = "white") +
  geom_point(data = cnp_med, aes(x = LON, y = LAT), alpha = 0.5, shape = 21) +
  expand_limits(x = world_map$long, y = world_map$lat) +
  viridis::scale_fill_viridis() +
  coord_equal(xlim = c(-10, 30), ylim = c(35, 62.5))

beta_country %>%
  ggplot() +
  geom_map(aes(map_id = region, fill = `log10(DOC_med)`), map = world_map, col = "white") +
  geom_point(data = cnp_med, aes(x = LON, y = LAT), alpha = 0.5, shape = 21) +
  expand_limits(x = world_map$long, y = world_map$lat) +
  viridis::scale_fill_viridis() +
  coord_equal(xlim = c(-10, 30), ylim = c(35, 62.5))

beta_country %>%
  ggplot() +
  geom_map(aes(map_id = region, fill = `log10(P_med)`), map = world_map, col = "white") +
  geom_point(data = cnp_med, aes(x = LON, y = LAT), alpha = 0.5, shape = 21) +
  expand_limits(x = world_map$long, y = world_map$lat) +
  viridis::scale_fill_viridis() +
  coord_equal(xlim = c(-10, 30), ylim = c(35, 62.5))

```


# locations of sites where country is NA

When assigning sites to countries using the function `maps::map.where()` `r cnp %>% filter(is.na(country)) %>% pull(SITE_ID) %>% unique() %>% length()` sites had no country designation. The following map shows the locations of the sites which we not assigned to any country. These were excluded from the analysis. However, these could be manually assigned a country in the future.

```{r}
# where are NA locations?
cnp %>% filter(is.na(country)) %>% 
  ggplot() +
  geom_map(data = world_map, map = world_map,
           aes(long, lat, map_id = region),
           color = "white", fill = "lightgray") +
  geom_point(mapping = aes(x = LON, y = LAT)) +
  coord_equal(ratio = 1, xlim = c(-80, 15), ylim = c(30, 60))
```

