---
title: "spatial clustering for CNP sites"
author: "David Nguyen"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(dbscan)
library(maps)

theme_set(theme_bw())
```

# Overview

To "objectively" decide how to subset the CNP data set I use hierarchical density-based spatial clustering of applications with noise (HDBSCAN). 

DBSCAN forms clusters based on "reachability" which allows for determination of odd shaped clusters. The hierarchical variant allows for clusters that have different densities.

I used great circle earth distance (`fields::rdist.earth()`) to cluster the sampling sites. I chose 50 points as the minimum cluster size.

```{r}
# load CNP data
cnp <- read_csv("https://raw.githubusercontent.com/swulfing/STOICH.Aim1/main/Data/ALL_CNP_April2022.csv")

# correcting DOC for LAGOS sites seems to make sense
# probably was a unit error
cnp <- cnp %>% mutate(DOC = ifelse(log10(DOC) < -1 & str_detect(SITE_ID, "LAGOS"),
                                   DOC * 10^3,
                                   DOC))
# get country from LAT LON
cnp <- 
  cnp %>% 
  mutate(country = map.where(database = "world", x = LON, y = LAT))


# get median of each SITE_ID over time
cnp_med <- 
  cnp %>% #filter(country %in% country_names) %>%
  # group_by(SITE_ID) %>%
  group_by(LAT, LON) %>%
  summarise(country = unique(country), 
            LAT = unique(LAT),
            LON = unique(LON),
            ECO_TYPE = unique(ECO_TYPE),
            N_med = median(NO3.as.N, na.rm = TRUE),
            P_med = median(PO4.as.P, na.rm = TRUE),
            DOC_med = median(DOC, na.rm = TRUE))
```

```{r eval = FALSE}
# calculate
D = fields::rdist.earth(x1 = cbind(cnp_med$LON, cnp_med$LAT))
# use hdbscan clustering
minPts <- 50
system.time(clust <-  hdbscan(x = D, minPts = minPts))
# save hdclust object
saveRDS(clust, file = "models/hdbscan_clusters.rds")
```

```{r}
clust <- readRDS("models/hdbscan_clusters.rds")
clust
```

```{r}
cluster_names <- c("Noise", "Baltics", "Germany", 
                   "Spain", "France", "Wales", "Scotland","North America"
                   )
cnp_med$cluster <- factor(clust$cluster, labels = cluster_names)
# cnp_med$cluster[cnp_med$cluster == 0] <- NA
# saveRDS(cnp_med, file = "data/cnp_median_clustered.rds")
```

```{r}
world_map <- map_data("world")

ggplot() +
  geom_map(aes(map_id = region), col = "grey", fill = "white", 
           data = world_map, map = world_map) +
  geom_point(data = cnp_med, aes(x = LON, y = LAT, col = cluster), shape = 21, alpha = 0.5) +
  expand_limits(x = c(min(cnp_med$LON), max(cnp_med$LON)), 
                      y = c(min(cnp_med$LAT), max(cnp_med$LAT))) +
  coord_equal(ratio = 1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, shape = 16)))
```

```{r}
cnp_med %>%
  group_by(cluster) %>%
  summarise(n_sites = n(),
            prop_lake = sum(ECO_TYPE == "Lake") / n(),
            prob_river = 1 - prop_lake) %>%
  knitr::kable(caption = "ECO_TYPE by each cluster",
               digits = 2)

```

